### 2024-11-04 informe/acl2023.tex
feat: agregado informe

\pdfoutput=1
\documentclass[11pt]{article}

\usepackage[table,xcdraw]{xcolor}
% \usepackage{ACL2023}
\usepackage{amsmath}
\usepackage{times}
\usepackage{latexsym}
\usepackage[T1]{fontenc}

\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{inconsolata}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{float}
\usepackage{subfig}
\usepackage{booktabs}
\usepackage{url}

%%%%%%%%%%%%%%%%%%% TITLE AND AUTHORS %%%%%%%%%%%%%%%%%%%

\title{IIC3633  Análisis \& Baselines proyecto de Recomendación multimodal de juegos de mesa}

\author{\normalfont 
Eduardo Salinas | 21624453 | \texttt{esalinasbarros@uc.cl} \\
Nicolás Gutiérrez | 20203772 | \texttt{njgutierrez@uc.cl} \\
Alfonso Badilla | 20640854 | \texttt{alfonso.badilla@uc.cl} \\
% comment line below if only 3 team members
%Author 4 | SCIPER 4 | \texttt{author4@epfl.ch} \\
}

%%%%%%%%%%%%%%%%%%% PROPOSAL %%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle
\newpage

\section{Descripción del problema y justificación}
Actualmente, el mercado de los juegos de mesa esta experimentando un crecimiento significativo, lo que implica una gran cantidad de juegos nuevos en el mercado cada día con una alta variedad de generas, temas y niveles de complejidad. Aunque existen clasificadores generales con reseñas, esto no toma en consideración los gustos personales.

Un sistema recomendador que sea capaz de recomendar juegos según fotos, descripciones sobre aspectos del juego, lenguaje natural e historial de gustos le da un aspecto de personalización al problema, mejorando la solución significativamente y potenciando la toma de decisiones a la hora de elegir un juego de mesa que encaje con lo esperado por el usuario. Además, dado el gran crecimiento del mercado de juegos un sistema como el explicado, sera cada vez mas utilizado y explotado para elegir juegos de calidad.

\section{Objetivos del proyecto}
El objetivo del proyecto es crear un sistema recomendador de juegos de mesa, que sea capaz de hacer las recomendaciones en base a imágenes y características descriptivas (metadata) utilizando filtrado colaborativo. La idea es que el sistema sea multimodal y que pueda recibir diferentes formatos de datos de entrada para dar recomendaciones validas que al usuario le gustaran. Para esto, se utilizaran métricas de testeo de sistemas recomendadores como precisión, novelty y diversidad.

\section{Analisis descriptivo de los datos}
El dataset que se utilizara para este proyecto consta de 10 archivos diferentes, cada uno con información diferente sobre juegos de mesa. El punto de entrada sera el archivo \texttt{user\_ratings.csv} el cual consiste de 19 millones de filas. Por temas de procesamiento limitado de Google Collab, se utilizaran 50.000 filas para el entrenamiento  inicial. A continuación ser presenta un desglose de los diferentes archivos a utilizar:

\begin{itemize}
    \item \texttt{games.csv}: Este archivo contiene 22 atributos sobre los juegos calificados y contiene 22.000 juegos diferentes. Entre los atributos tenemos aspectos como la descripción del juego, que se podrían evaluar usando modelos de lenguaje, el año de publicación (que distribuye como se muestra en la figura \ref{fig:publishYear}), el rating promedio recibido, entre otros.
    \item \texttt{ratings\_distribution.csv}: Contiene el total de ratings en cada juego de mesa. En general estos ratings distribuyen como se ve en la figura \ref{fig:distribucionRatings}.
    \item \texttt{temes.csv}: Contiene las temáticas de cada juego como un flag binario. Las 50 más comunes se pueden ver en el gráfico de la figura \ref{fig:tematicasComunes}.
    \item \texttt{mechanics.csv}: Contiene las mecánicas del juego como un flag binario. Las 50 más comunes se pueden ver en el gráfico de la figura \ref{fig:mecanicasComunes}.
    \item \texttt{subcategories.csv}: Contiene las subcategorías de cada juego en formato flag binario.
    \item \texttt{artists\_reduced.csv}: Contiene la información sobre que artista creo cierto juego en formato flag binario. Solo se consideran artistas con mas de 3 juegos creados.
    \item \texttt{designer\_reduced.csv}: Contiene la información sobre que diseñador diseño cierto juego en formato flag binario. Solo se consideran diseñadores con mas de 3 juegos creados.
    \item \texttt{publishers\_reduced.csv}: Contiene la información de las empresas que venden dichos juegos en formato flag binario. Solo se consideran las empresas que veden mas de 3 juegos.

\end{itemize}

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{publishYear.png}
    \caption{Juegos por año de publicación}
    \label{fig:publishYear}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\linewidth]{distribucionRatings.png}
    \caption{Distribución de ratings de juegos}
    \label{fig:distribucionRatings}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\linewidth]{tematicasComunes.png}
    \caption{Temáticas más comunes en el dataset}
    \label{fig:tematicasComunes}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\linewidth]{mecanicasComunes.png}
    \caption{Mecánicas más comunes en el dataset}
    \label{fig:mecanicasComunes}
\end{figure}

\section{Implementación de modelos de referencia} 
Para tener una idea basica de las metricas de recomendación actuales, entrenamos 4 modelos y obtuvimos las siguientes metricas.
\begin{table}[h!]
\centering
\begin{tabular}{|l|c|c|c|c|}
\hline
\textbf{Modelo}    & \textbf{MAE}   & \textbf{RMSE}  & \textbf{MAP}     & \textbf{NDCG}        \\ \hline
ItemKNN            & 7.1102         & 7.1102         & 1.15e-4          & 5.873e-5             \\ \hline
SVD                & 7.2539         & 7.2539         & 1.54e-4          & 5.873e-5             \\ \hline
MostPopular        & -              & -              & 0.077            & 0.057                \\ \hline
Random             & 2.827          & 3.45           & -                & -                    \\ \hline
\end{tabular}
\caption{Resultados de diferentes modelos de recomendación}
\end{table}
\newline
Se pueden observar errores relativamente altos para los modelos ItemKNN y SVD y metricas de precision bastante bajas. Los otros algoritmos tampoco están mostrando un desempeño muy alto, lo que hace referencia a una falta de calidad en el dataset. Dicho esto, si bien hay una falta de metricas, se puede observar que MostPopular y Random funcionan mejor que ItemKNN y SVD.

\section{Planificación para la etapa Midterm} 
El trabajo que realizaremos se espera que se distribuya en dos partes:
\begin{enumerate}
    \item \textbf{Aprovechar metadata cualitativo:} Primero se hará una comparación de los modelos de baseline con uno que incorpore el metadata de las tablas adicionales del dataset entregado, y se verá si se logra ganancia de precisión y otras métricas considerando ese metadata. Se espera tener esto listo entre la segunda y tercera entrega, con al menos un avance en la segunda, y se debería evaluar en función de la robustez de la utilización del metadata relevante para entrenar modelos, es decir si se utilizó el metadata de una cantidad de formas suficientes para entrenar un modelo suficientemente bueno.
    \item \textbf{Aprovechar imágenes y descripción de los juegos:} Luego se repetirá este experimento pero agregando además como metadata las imágenes y la descripción relacionados con los respectivos juegos de mesa, considerando también performance en tiempo del modelo, para revisar si se obtiene ganancia suficiente de considerar estos datos dentro del metadata cualitativo del modelo (se agragecería feedback especialmente en la factibilidad y forma de hacer esta parte del trabajo). Si se realizara, se haría para la tercera entrega, y se debería evaluar en función de la profundidad de la investigación relacionada con la aplicación de estas formas de metadata.
\end{enumerate}

\section{Bibliografía relevante}

\begin{itemize}
    \item En el artículo "Addressing the user cold start with cross-domain collaborative filtering: exploiting item metadata in matrix factorization" \url{https://link.springer.com/article/10.1007/s11257-018-9217-6} se utiliza filtrado colaborativo y se busca reducir el efecto del "cold start" de un usuario nuevo utilizando el metadata de los ítems de la plataforma.
    \item En el artículo "Using geospatial metadata to boost collaborative filtering" \url{https://dl.acm.org/doi/abs/10.1145/2507157.2507201} se utiliza metadata de espacio para recomendar ítems del tipo fotos junto con filtrado colaborativo.
\end{itemize}

\end{document}

### 2024-11-09 informe/acl2023.tex
fix: FUCK LATEX (arreglado)
feat: respaldo e1 + inicio informe e2

@@
-\pdfoutput=1
+%\pdfoutput=1
 \documentclass[11pt]{article}
@@
 \usepackage{amsmath}
-\usepackage{times}
 \usepackage{latexsym}
@@
 \usepackage[utf8]{inputenc}
-\usepackage{microtype}
-\usepackage{inconsolata}
 \usepackage{fancyhdr}
@@
 \usepackage{url}
+\usepackage{comment}
 
@@
 
-\title{IIC3633  Análisis \& Baselines proyecto de Recomendación multimodal de juegos de mesa}
+\title{IIC3633 Aplicación de recomendación grupal de juegos de mesa utilizando metadata}
 
@@
 
-\section{Descripción del problema y justificación}
-Actualmente, el mercado de los juegos de mesa esta experimentando un crecimiento significativo, lo que implica una gran cantidad de juegos nuevos en el mercado cada día con una alta variedad de generas, temas y niveles de complejidad.
-Aunque existen clasificadores generales con reseñas, esto no toma en consideración los gustos personales.
+\section{Pivoteo en la idea original del proyecto}
 
-Un sistema recomendador que sea capaz de recomendar juegos según fotos, descripciones sobre aspectos del juego, lenguaje natural e historial de gustos le da un aspecto de personalización al problema, mejorando la solución significativamente y potenciando la toma de decisiones a la hora de elegir un juego de mesa que encaje con lo esperado por el usuario.
-Además, dado el gran crecimiento del mercado de juegos un sistema como el explicado, sera cada vez mas utilizado y explotado para elegir juegos de calidad.
+El proyecto originalmente consistía en un sistema recomendador de juegos de mesa, que fuera capaz de hacer recomendaciones en base a imágenes y características descriptivas (metadata).
+La idea era que el sistema fuera multimodal y que pudiera recibir diferentes formatos de datos de entrada para dar recomendaciones válidas que al usuario le gustaran.
 
-\section{Objetivos del proyecto}
-El objetivo del proyecto es crear un sistema recomendador de juegos de mesa, que sea capaz de hacer las recomendaciones en base a imágenes y características descriptivas (metadata) utilizando filtrado colaborativo.
-La idea es que el sistema sea multimodal y que pueda recibir diferentes formatos de datos de entrada para dar recomendaciones validas que al usuario le gustaran.
-Para esto, se utilizaran métricas de testeo de sistemas recomendadores como precisión, novelty y diversidad.
+Nosotros decidimos cambiar la idea original del proyecto por una que consideramos que podría llegar a ser más poderosa y que a nivel básico debería devolver resultados pronto.
+La nueva idea consiste en un sistema recomendador de juegos de mesa, que sea capaz de hacer recomendaciones para grupos en base a características descriptivas (metadata).
+La idea es que el sistema sea multimodal y que pueda recibir diferentes formatos de datos de entrada para dar recomendaciones válidas que al grupo de usuarios le gusten.
 
-\section{Analisis descriptivo de los datos}
-El dataset que se utilizara para este proyecto consta de 10 archivos diferentes, cada uno con información diferente sobre juegos de mesa.
-El punto de entrada sera el archivo \texttt{user\_ratings.csv} el cual consiste de 19 millones de filas.
-Por temas de procesamiento limitado de Google Collab, se utilizaran 50.000 filas para el entrenamiento  inicial.
-A continuación ser presenta un desglose de los diferentes archivos a utilizar:
+\section{Progreso en el desarrollo de la solución}
 
+Para esta entrega, nosotros hemos entrenado un recomendador sencillo utilizando el algoritmo de filtrado colaborativo.
+Para esto, utilizamos el dataset de BoardGameGeek que contiene información sobre juegos de mesa y ratings de usuarios.
+El dataset contiene 10 archivos diferentes, cada uno con información diferente sobre juegos de mesa.
+El punto de entrada fue el archivo \texttt{user\_ratings.csv} el cual consiste de 19 millones de filas.
+Por temas de procesamiento limitado de Google Collab (y más tarde de nuestro computador local), se utilizaron 50.000 filas para el entrenamiento inicial.
+
+% Actualizar esto cuando entrenemos el recomendador que use metadata!!!
+
+\section{Experimentación realizada y evaluación intemredia}
+A modo de experimentación se entrenaron 4 modelos de recomendación diferentes: ItemKNN, SVD, MostPopular y Random que luego se usarán como baseline.
+De estos, se utilizó ItemKNN para hacer una recomendación de juegos para grupos utilizando la ponderación utilitaria-aditiva (equivalente al promedio) la cual se usará como baseline para esta entrega.
+
+\section{Analisis preliminar de los resultados}
+El recomendador que utiliza metadata lo comparamos con el de ItemKNN para varios grupos de usuarios de 5 integrantes. Para ellos encontramos los siguientes resultados:
+
+% Reemplazar las tablas cuando tengamos los resultados de los modelos.
+
+\begin{comment}
 \begin{itemize}
@@
 
-\section{Implementación de modelos de referencia} 
-Para tener una idea basica de las metricas de recomendación actuales, entrenamos 4 modelos y obtuvimos las siguientes metricas.
-\begin{table}[h!]
-\centering
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-\textbf{Modelo}    & \textbf{MAE}   & \textbf{RMSE}  & \textbf{MAP}     & \textbf{NDCG}        \\ \hline
-ItemKNN            & 7.1102         & 7.1102         & 1.15e-4          & 5.873e-5             \\ \hline
-SVD                & 7.2539         & 7.2539         & 1.54e-4          & 5.873e-5             \\ \hline
-MostPopular        & -              & -              & 0.077            & 0.057                \\ \hline
-Random             & 2.827          & 3.45           & -                & -                    \\ \hline
-\end{tabular}
-\caption{Resultados de diferentes modelos de recomendación}
-\end{table}
-\newline
-Se pueden observar errores relativamente altos para los modelos ItemKNN y SVD y metricas de precision bastante bajas.
-Los otros algoritmos tampoco están mostrando un desempeño muy alto, lo que hace referencia a una falta de calidad en el dataset.
-Dicho esto, si bien hay una falta de metricas, se puede observar que MostPopular y Random funcionan mejor que ItemKNN y SVD.
+\end{comment}
 
-\section{Planificación para la etapa Midterm} 
-El trabajo que realizaremos se espera que se distribuya en dos partes:
-\begin{enumerate}
-    \item \textbf{Aprovechar metadata cualitativo:} Primero se hará una comparación de los modelos de baseline con uno que incorpore el metadata de las tablas adicionales del dataset entregado, y se verá si se logra ganancia de precisión y otras métricas considerando ese metadata.
-Se espera tener esto listo entre la segunda y tercera entrega, con al menos un avance en la segunda, y se debería evaluar en función de la robustez de la utilización del metadata relevante para entrenar modelos, es decir si se utilizó el metadata de una cantidad de formas suficientes para entrenar un modelo suficientemente bueno.
-    \item \textbf{Aprovechar imágenes y descripción de los juegos:} Luego se repetirá este experimento pero agregando además como metadata las imágenes y la descripción relacionados con los respectivos juegos de mesa, considerando también performance en tiempo del modelo, para revisar si se obtiene ganancia suficiente de considerar estos datos dentro del metadata cualitativo del modelo (se agragecería feedback especialmente en la factibilidad y forma de hacer esta parte del trabajo).
-Si se realizara, se haría para la tercera entrega, y se debería evaluar en función de la profundidad de la investigación relacionada con la aplicación de estas formas de metadata.
-\end{enumerate}
+\section{Problemas identificados durante el proceso}
 
-\section{Bibliografía relevante}
+Previo al pivoteo, los principales problemas que enfrentamos fueron la falta de recursos computacionales para entrenar modelos de recomendación que utilizan imágenes y la falta de experiencia en el manejo de datos de imágenes.
+% Creo que teníamos un problema más pero no me acuerdo cuál era.
 
+Luego del pivoteo, los problemas que enfrentamos fueron... % Agregar los problemas que enfrentamos luego del pivoteo.
+
+\section{Revisión del plan propuesto en la etapa anterior y justificación de ajustes}
+
+En general, el pivoteo fue\dots %Qué tal nos fue con el pivoteo?
+
+Para la próxima entrega, planeamos hacer que cada usuario pueda elegir que cada una de las metadatas que le corresponden tenga un peso distinto en la recomendación.
+Esto lo haremos \dots %Dejar alguna idea de cómo lo haremos escrita.
+
+\section{Anexo: actualización de propuesta}
 \begin{itemize}
-    \item En el artículo "Addressing the user cold start with cross-domain collaborative filtering: exploiting item metadata in matrix factorization" \url{https://link.springer.com/article/10.1007/s11257-018-9217-6} se utiliza filtrado colaborativo y se busca reducir el efecto del "cold start" de un usuario nuevo utilizando el metadata de los ítems de la plataforma.
-    \item En el artículo "Using geospatial metadata to boost collaborative filtering" \url{https://dl.acm.org/doi/abs/10.1145/2507157.2507201} se utiliza metadata de espacio para recomendar ítems del tipo fotos junto con filtrado colaborativo.
+    \item Deben agregar unos 3 baselines de métodos para recomendar grupos
+    \item Deben actualizar la metodología de evaluación indicando métricas especificas para recomendación a grupos, no pueden hacer la misma evaluación que recomendar a individuos
 \end{itemize}

### 2024-11-09 informe/respaldo entrega 1/respaldo.tex
feat: respaldo e1 + inicio informe e2

%\pdfoutput=1
\documentclass[11pt]{article}

\usepackage[table,xcdraw]{xcolor}
% \usepackage{ACL2023}
\usepackage{amsmath}
\usepackage{latexsym}
\usepackage[T1]{fontenc}

\usepackage[utf8]{inputenc}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{float}
\usepackage{subfig}
\usepackage{booktabs}
\usepackage{url}

\begin{document}
\section{Descripción del problema y justificación}
Actualmente, el mercado de los juegos de mesa esta experimentando un crecimiento significativo, lo que implica una gran cantidad de juegos nuevos en el mercado cada día con una alta variedad de generas, temas y niveles de complejidad. Aunque existen clasificadores generales con reseñas, esto no toma en consideración los gustos personales.

Un sistema recomendador que sea capaz de recomendar juegos según fotos, descripciones sobre aspectos del juego, lenguaje natural e historial de gustos le da un aspecto de personalización al problema, mejorando la solución significativamente y potenciando la toma de decisiones a la hora de elegir un juego de mesa que encaje con lo esperado por el usuario. Además, dado el gran crecimiento del mercado de juegos un sistema como el explicado, sera cada vez mas utilizado y explotado para elegir juegos de calidad.

\section{Objetivos del proyecto}
El objetivo del proyecto es crear un sistema recomendador de juegos de mesa, que sea capaz de hacer las recomendaciones en base a imágenes y características descriptivas (metadata) utilizando filtrado colaborativo. La idea es que el sistema sea multimodal y que pueda recibir diferentes formatos de datos de entrada para dar recomendaciones validas que al usuario le gustaran. Para esto, se utilizaran métricas de testeo de sistemas recomendadores como precisión, novelty y diversidad.

\section{Analisis descriptivo de los datos}
El dataset que se utilizara para este proyecto consta de 10 archivos diferentes, cada uno con información diferente sobre juegos de mesa. El punto de entrada sera el archivo \texttt{user\_ratings.csv} el cual consiste de 19 millones de filas. Por temas de procesamiento limitado de Google Collab, se utilizaran 50.000 filas para el entrenamiento  inicial. A continuación ser presenta un desglose de los diferentes archivos a utilizar:

\begin{itemize}
    \item \texttt{games.csv}: Este archivo contiene 22 atributos sobre los juegos calificados y contiene 22.000 juegos diferentes. Entre los atributos tenemos aspectos como la descripción del juego, que se podrían evaluar usando modelos de lenguaje, el año de publicación (que distribuye como se muestra en la figura \ref{fig:publishYear}), el rating promedio recibido, entre otros.
    \item \texttt{ratings\_distribution.csv}: Contiene el total de ratings en cada juego de mesa. En general estos ratings distribuyen como se ve en la figura \ref{fig:distribucionRatings}.
    \item \texttt{temes.csv}: Contiene las temáticas de cada juego como un flag binario. Las 50 más comunes se pueden ver en el gráfico de la figura \ref{fig:tematicasComunes}.
    \item \texttt{mechanics.csv}: Contiene las mecánicas del juego como un flag binario. Las 50 más comunes se pueden ver en el gráfico de la figura \ref{fig:mecanicasComunes}.
    \item \texttt{subcategories.csv}: Contiene las subcategorías de cada juego en formato flag binario.
    \item \texttt{artists\_reduced.csv}: Contiene la información sobre que artista creo cierto juego en formato flag binario. Solo se consideran artistas con mas de 3 juegos creados.
    \item \texttt{designer\_reduced.csv}: Contiene la información sobre que diseñador diseño cierto juego en formato flag binario. Solo se consideran diseñadores con mas de 3 juegos creados.
    \item \texttt{publishers\_reduced.csv}: Contiene la información de las empresas que venden dichos juegos en formato flag binario. Solo se consideran las empresas que veden mas de 3 juegos.

\end{itemize}

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{publishYear.png}
    \caption{Juegos por año de publicación}
    \label{fig:publishYear}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\linewidth]{distribucionRatings.png}
    \caption{Distribución de ratings de juegos}
    \label{fig:distribucionRatings}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\linewidth]{tematicasComunes.png}
    \caption{Temáticas más comunes en el dataset}
    \label{fig:tematicasComunes}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\linewidth]{mecanicasComunes.png}
    \caption{Mecánicas más comunes en el dataset}
    \label{fig:mecanicasComunes}
\end{figure}

\section{Implementación de modelos de referencia} 
Para tener una idea basica de las metricas de recomendación actuales, entrenamos 4 modelos y obtuvimos las siguientes metricas.
\begin{table}[h!]
\centering
\begin{tabular}{|l|c|c|c|c|}
\hline
\textbf{Modelo}    & \textbf{MAE}   & \textbf{RMSE}  & \textbf{MAP}     & \textbf{NDCG}        \\ \hline
ItemKNN            & 7.1102         & 7.1102         & 1.15e-4          & 5.873e-5             \\ \hline
SVD                & 7.2539         & 7.2539         & 1.54e-4          & 5.873e-5             \\ \hline
MostPopular        & -              & -              & 0.077            & 0.057                \\ \hline
Random             & 2.827          & 3.45           & -                & -                    \\ \hline
\end{tabular}
\caption{Resultados de diferentes modelos de recomendación}
\end{table}
\newline
Se pueden observar errores relativamente altos para los modelos ItemKNN y SVD y metricas de precision bastante bajas. Los otros algoritmos tampoco están mostrando un desempeño muy alto, lo que hace referencia a una falta de calidad en el dataset. Dicho esto, si bien hay una falta de metricas, se puede observar que MostPopular y Random funcionan mejor que ItemKNN y SVD.

\section{Planificación para la etapa Midterm} 
El trabajo que realizaremos se espera que se distribuya en dos partes:
\begin{enumerate}
    \item \textbf{Aprovechar metadata cualitativo:} Primero se hará una comparación de los modelos de baseline con uno que incorpore el metadata de las tablas adicionales del dataset entregado, y se verá si se logra ganancia de precisión y otras métricas considerando ese metadata. Se espera tener esto listo entre la segunda y tercera entrega, con al menos un avance en la segunda, y se debería evaluar en función de la robustez de la utilización del metadata relevante para entrenar modelos, es decir si se utilizó el metadata de una cantidad de formas suficientes para entrenar un modelo suficientemente bueno.
    \item \textbf{Aprovechar imágenes y descripción de los juegos:} Luego se repetirá este experimento pero agregando además como metadata las imágenes y la descripción relacionados con los respectivos juegos de mesa, considerando también performance en tiempo del modelo, para revisar si se obtiene ganancia suficiente de considerar estos datos dentro del metadata cualitativo del modelo (se agragecería feedback especialmente en la factibilidad y forma de hacer esta parte del trabajo). Si se realizara, se haría para la tercera entrega, y se debería evaluar en función de la profundidad de la investigación relacionada con la aplicación de estas formas de metadata.
\end{enumerate}

\section{Bibliografía relevante}

\begin{itemize}
    \item En el artículo "Addressing the user cold start with cross-domain collaborative filtering: exploiting item metadata in matrix factorization" \url{https://link.springer.com/article/10.1007/s11257-018-9217-6} se utiliza filtrado colaborativo y se busca reducir el efecto del "cold start" de un usuario nuevo utilizando el metadata de los ítems de la plataforma.
    \item En el artículo "Using geospatial metadata to boost collaborative filtering" \url{https://dl.acm.org/doi/abs/10.1145/2507157.2507201} se utiliza metadata de espacio para recomendar ítems del tipo fotos junto con filtrado colaborativo.
\end{itemize}

\end{document}

### 2024-11-11 informe/acl2023.tex
Metadata recommendation algorithm

@@
 A modo de experimentación se entrenaron 4 modelos de recomendación diferentes: ItemKNN, SVD, MostPopular y Random que luego se usarán como baseline. 
-De estos, se utilizó ItemKNN para hacer una recomendación de juegos para grupos utilizando la ponderación utilitaria-aditiva (equivalente al promedio) la cual se usará como baseline para esta entrega.
+De estos, se utilizó ItemKNN para hacer una recomendación de juegos para grupos utilizando la ponderación utilitaria-aditiva (equivalente al promedio) 
+la cual se usará como baseline para esta entrega. Ademas, se integro un modelo de la libreria \texttt{fastFM} para crear un modelo capaz de recibir metadata y dar recomendaciones.
+Para este caso particular, se entreno el modelo agregando features a items los cuales consistian en las categorias en las que se clasificaba el item a recomendar. Las features que se utilizaron 
+consistian en la categoria del juego y la mecanica de este. Estos valores se convirtieron desde un flag binario a una representación vectorial como embeddings.
 
@@
 El recomendador que utiliza metadata lo comparamos con el de ItemKNN para varios grupos de usuarios de 5 integrantes. Para ellos encontramos los siguientes resultados:
+
+\begin{table}[H]
+    \centering
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        \textbf{Modelo} & \textbf{Recall@10} & \textbf{Precision@10} & \textbf{NDCG@10} & \textbf{RMSE} \\ \hline
+        ItemKNN & - & 0.000386 & 0.000128 & 7.13 \\ \hline
+        fastFM \& metadata & 0.107 & 0.0056 & - & - \\ \hline
+    \end{tabular}
+    \caption{Resultados preliminares de recomendaciones para grupos}
+    \label{tab:resultados_preliminares}
+\end{table}
+
+Si bien no todas las metricas estan incluidas, se puede observar que el modelo que utiliza metadata tiene un mejor recall y precision que el modelo basado en itemKNN.
 
@@
 
-Luego del pivoteo, los problemas que enfrentamos fueron... % Agregar los problemas que enfrentamos luego del pivoteo.
+Luego del pivoteo, los problemas que enfrentamos fueron:
+\begin{itemize}
+    \item Falta de obtención de metricas para comparar modelos que usan metadata.
+    \item Poca flexibilidad de los modelos que utilizan metadata para recibir diferentes formatos de entrada.
+\end{itemize}
 

### 2024-11-11 informe/acl2023.tex
Update LaTeX
Agregado en el párrafo que dice "Luego del pivoteo, ..." el limite del sample de 5000 reseñas

--- 2024-11-09 informe/acl2023.tex
@@
 
-Luego del pivoteo, los problemas que enfrentamos fueron... % Agregar los problemas que enfrentamos luego del pivoteo.
+Luego del pivoteo, los problemas que enfrentamos fueron, por una parte, la capacidad limitada de cómputo de Google Colab, donde solo podíamos generar grupos con pequeñas muestras, no de una cantidad que sea representativa del dataset en general, hasta el momento, trabajamos con un sample de 5000 reseñas.
+...
+% Agregar los problemas que enfrentamos luego del pivoteo.
 

### 2024-11-11 informe/acl2023.tex
feat: actualizado informe para entrega 2

--- 2024-11-11 informe/acl2023.tex
@@
 \usepackage[table,xcdraw]{xcolor}
+\usepackage[a4paper, margin=1in]{geometry}
 % \usepackage{ACL2023}
@@
 \usepackage{comment}
+\usepackage{hyperref}
 
@@
 
-\section{Pivoteo en la idea original del proyecto}
-
-El proyecto originalmente consistía en un sistema recomendador de juegos de mesa, que fuera capaz de hacer recomendaciones en base a imágenes y características descriptivas (metadata).
-La idea era que el sistema fuera multimodal y que pudiera recibir diferentes formatos de datos de entrada para dar recomendaciones válidas que al usuario le gustaran.
-
-Nosotros decidimos cambiar la idea original del proyecto por una que consideramos que podría llegar a ser más poderosa y que a nivel básico debería devolver resultados pronto.
-La nueva idea consiste en un sistema recomendador de juegos de mesa, que sea capaz de hacer recomendaciones para grupos en base a características descriptivas (metadata).
-La idea es que el sistema sea multimodal y que pueda recibir diferentes formatos de datos de entrada para dar recomendaciones válidas que al grupo de usuarios le gusten.
-
 \section{Progreso en el desarrollo de la solución}
 
-Para esta entrega, nosotros hemos entrenado un recomendador sencillo utilizando el algoritmo de filtrado colaborativo.
+Para esta entrega hicimos el pivoteo visto en la sección 6, primero hemos entrenado varios recomendadores sencillos de baseline utilizando filtrado colaborativo, most popular, random entre otros.
 Para esto, utilizamos el dataset de BoardGameGeek que contiene información sobre juegos de mesa y ratings de usuarios.
@@
 
-% Actualizar esto cuando entrenemos el recomendador que use metadata!!!
+Luego, hicimos un recomendador más poderoso para este dataset, que es el que utiliza metadata.
+Para esto, utilizamos las categorías y mecánicas de los juegos mostrados en los otros archivos como metadata.
+Para esto, se utilizó la librería \texttt{fastFM} que permite entrenar modelos de factorización de matrices rápidamente.
+Se entrenó un modelo que recibe las categorías y mecánicas de los juegos y entrega recomendaciones para grupos de usuarios.
+
+Además, implementamos en nuestro dataset un generador de grupos que se encarga de generar grupos de usuarios de tamaño fijo.
+Este generador de grupos se copió del trabajo mostrado en el paper \textit{Group Recommendation with Fast Matrix Factorization} de \textit{Yuan et al.} recuperado de su repositorio público en \href{https://github.com/barnap/group-recommenders-offline-evaluation}{github.com}.
+
+No se logró implementar la recomendación a los grupos generados utilizando el modelo con metadata, debido a que por problemas de tiempo no se pudo hacer el ajuste necesario para que el modelo reciba los datos de entrada en el formato correcto, dado que nuestro modelo con metadata no entrega los reviews predichos, sino que entrega un orden de preferencia de los juegos (se tiene que adaptar la preferencia a un formato de predicción de rating para que funcione con las estrategias de agregación del repositorio de Group Recommendation).
+
+En general la dirección en la que llevamos este trabajo resultó en que terminará siendo una expansión del trabajo de Yuan et al.
+en el sentido de que será una experimentación de qué tal funciona con un modelo con metadata.
 
@@
 A modo de experimentación se entrenaron 4 modelos de recomendación diferentes: ItemKNN, SVD, MostPopular y Random que luego se usarán como baseline.
-De estos, se utilizó ItemKNN para hacer una recomendación de juegos para grupos utilizando la ponderación utilitaria-aditiva (equivalente al promedio) 
-la cual se usará como baseline para esta entrega. Ademas, se integro un modelo de la libreria \texttt{fastFM} para crear un modelo capaz de recibir metadata y dar recomendaciones.
+Además, se integrá un modelo de la librería \texttt{fastFM} para crear un modelo capaz de recibir metadata y dar recomendaciones.
 Para este caso particular, se entreno el modelo agregando features a items los cuales consistian en las categorias en las que se clasificaba el item a recomendar. Las features que se utilizaron 
 consistian en la categoria del juego y la mecanica de este. Estos valores se convirtieron desde un flag binario a una representación vectorial como embeddings.
+Se espera a futuro comparar el modelo generado con uno creado utilizando ItemKNN para ver si el modelo con metadata es capaz de dar mejores recomendaciones.
+Además, queremos tratar de mejorar el modelo con metadata, revisando qué tan factible es que un usuario asigne mayor peso a ciertos embeddings que a otros, para que el modelo pueda dar recomendaciones más personalizadas para este usuario único dentro del grupo.
 
@@
 
-% Reemplazar las tablas cuando tengamos los resultados de los modelos.
+En la próxima entrega deberíamos hacer cierta variación de parámetros para asegurar que sea la mejor versión de nuestro modelo con metadata.
 
@@
 Previo al pivoteo, los principales problemas que enfrentamos fueron la falta de recursos computacionales para entrenar modelos de recomendación que utilizan imágenes y la falta de experiencia en el manejo de datos de imágenes.
-% Creo que teníamos un problema más pero no me acuerdo cuál era.
+Este problema decidimos solucionarlo cambiando la idea original del proyecto, a la cual esperamos acercarnos con recursos computacionales más poderosos.
 
@@
     \item Poca flexibilidad de los modelos que utilizan metadata para recibir diferentes formatos de entrada.
+    \item La capacidad limitada de cómputo de Google Colab, donde solo podíamos generar grupos con pequeñas muestras.
+Dado que la cantidad de 5000 grupos límite que encontramos mediante experimentación no es suficiente para hacer una evaluación robusta de los modelos, se decidió mover ese trabajo a un computador local, y se logró resolver el problema, generando en sólo 1 minuto un total de 50000 grupos los cuales se guardaron en un csv.
+    \item Se espera que a futuro usar estrategias más avanzadas de creación de grupos genere que se demore más tiempo en generar los grupos, por lo que se espera que este problema se vuelva a presentar en el futuro.
 \end{itemize}
@@
 
-En general, el pivoteo fue\dots %Qué tal nos fue con el pivoteo?
+En general, el pivoteo parece haber sido positivo, dado que gracias a este y el feedback del profesor se llegó a una idea de trabajo que busca expandir realmente el estado del arte, lo cual a efectos prácticos nos motivó y nos dio una base técnica sobre la cual construír más fácilmente.
 
-Para la próxima entrega, planeamos hacer que cada usuario pueda elegir que cada una de las metadatas que le corresponden tenga un peso distinto en la recomendación.
-Esto lo haremos \dots %Dejar alguna idea de cómo lo haremos escrita.
+Para la próxima entrega, planeamos terminar el recomendador grupal usando un método personalizado de simulación de ratings para poder usar las funciones de agregación del repositorio de Yuan et al.
 
-\section{Anexo: actualización de propuesta}
+También tendremos que solicitar feedback sobre la utilización de un repositorio público para la generación de grupos y las funciones de agregación y evaluación, ya que tenemos que compatibilizar su utilización con el código de honor.
+
+Además, veremos si se puede hacer que cada usuario pueda elegir que cada uno de los embeddings que le corresponden tenga un peso distinto en la recomendación.
+Esto lo revisaremos en detalle para la entrega 3, pero puede que no sea factible o se tenga que sacar de la planeación si nos falta tiempo.
+
+\newpage
+
+\section{Anexo: Pivoteo en la idea original del proyecto}
+
+El proyecto originalmente consistía en un sistema recomendador de juegos de mesa, que fuera capaz de hacer recomendaciones en base a imágenes y características descriptivas (metadata).
+La idea era que el sistema fuera multimodal y que pudiera recibir diferentes formatos de datos de entrada para dar recomendaciones válidas que al usuario le gustaran.
+
+Nosotros decidimos cambiar la idea original del proyecto por una que consideramos que podría llegar a ser más poderosa y que a nivel básico debería devolver resultados pronto.
+La nueva idea consiste en un sistema recomendador de juegos de mesa, que sea capaz de hacer recomendaciones para grupos en base a características descriptivas (metadata).
+La idea es que el sistema sea multimodal y que pueda recibir diferentes formatos de datos de entrada para dar recomendaciones válidas que al grupo de usuarios le gusten.
+
+\section{Anexo: Actualización de propuesta}
+Para hacer el pivoteo, el profesor nos pidió que hiciéramos una actualización de la propuesta a la cual debíamos agregar los siguientes puntos:
 \begin{itemize}
-    \item Deben agregar unos 3 baselines de métodos para recomendar grupos
-    \item Deben actualizar la metodología de evaluación indicando métricas especificas para recomendación a grupos, no pueden hacer la misma evaluación que recomendar a individuos
+    \item Deben agregar alrededor de 3 baselines de métodos para recomendar grupos.
+    \begin{enumerate}
+        \item Random es un baseline que nos permitirá saber si nuestro modelo es mejor que un modelo que recomienda juegos aleatorios. Este no debería ser difícil de superar.
+        \item Most popular directamente es un baseline que podría generarnos muchos problemas superar porque cuando se recomienda a un grupo, se tiende a recomendar el juego más popular para ese tipo de grupo, y para grupos generados aleatoriamente será difícil de mejorar.
+        \item ItemKNN usando la misma agregación que el modelo de metadata nos permitirá saber si el modelo de metadata mejoraría el resultado obtenido en el paper original.
+    \end{enumerate}
+    \item Deben actualizar la metodología de evaluación indicando métricas especificas para recomendación a grupos, no pueden hacer la misma evaluación que recomendar a individuos.
+    \begin{itemize}
+        \item Para la recomendación a grupos, se utilizarán las mismas métricas de evaluación que en el paper original, para verificar que nuestra expansión sea válida. Estas métricas son:
+        \begin{enumerate}
+            \item NDCGE, que mide la relevancia de los juegos recomendados de forma similar a NDCG para un recomendador normal.
+            \item DCG, que es NDCG sin normalizar.
+            \item BINARY, la métrica de evaluación binaria que se utiliza en el paper original.
+Esta consiste en revisar para cada recomendación, si es correcta, y luego medir la proporción de recomendaciones correctas.
+Luego, se aplica esta métrica según la política de agregación que se esté utilizando.
+            \item BASE, esta métrica en el paper parece ser para comparar con el modelo de recomendación base que se está utilizando.
+Esta métrica probablemente no la utilizaremos en el trabajo, excepto quizá para comparar con los modelos de most popular y random, si nos resulta útil.
+        \end{enumerate}
+    \end{itemize}
 \end{itemize}

### 2024-11-29 paper/paper.tex
feat: paper con titulo y abstract

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `all,proceedings,bibtex,sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigconf]{acmart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmlicensed}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}
%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Conference acronym 'XX]{Make sure to enter the correct
  conference title from your rights confirmation emai}{June 03--05,
  2018}{Woodstock, NY}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Board games group recomendation using metadata}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

\author{Eduardo Salinas}
\affiliation{
  \institution{Pontificia Universidad Católica}
  \city{Santiago}
  \country{Chile}}
\email{esalinasbarros@uc.cl}

\author{Alfonso Badilla}
\affiliation{
  \institution{Pontificia Universidad Católica}
  \city{Santiago}
  \country{Chile}}
\email{alfonso.badilla@uc.cl}

\author{Nicolás Gutiérrez}
\affiliation{
  \institution{Pontificia Universidad Católica}
  \city{Santiago}
  \country{Chile}}
\email{njgutierrez@uc.cl}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Trovato et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
    In the following paper, we propose a group recommendation system for the dataset on the website \href{https://boardgamegeek.com}{Board Game Geek}. The dataset contains information about board games, such as the number of players, the average time of a game, the complexity of the game, and the average rating of the game. 
    The recommendation system will be based on the K-means algorithm, which will group the games according to their characteristics. The system will recommend a group of games to a user based on the games they have rated. The recommendation system will be evaluated using the silhouette score, which will measure the quality of the clusters. 
    The results of the recommendation system will be compared with the results of a recommendation system based on the K-nearest neighbors algorithm and another algorithm described on the paper provided. %todo: explicar este algoritmo 
    The recommendation system will be implemented using the Python programming language and the scikit-learn library.
  % todo: mejorar abstract
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>00000000.0000000.0000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc[300]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc[100]{Do Not Use This Code~Generate the Correct Terms for Your Paper}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Do, Not, Us, This, Code, Put, the, Correct, Terms, for,
  Your, Paper}
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \includegraphics[width=\textwidth]{sampleteaser}
  \caption{Seattle Mariners at Spring Training, 2010.}
  \Description{Enjoying the baseball game from the third-base
  seats. Ichiro Suzuki preparing to bat.}
  \label{fig:teaser}
\end{teaserfigure}

\received{20 February 2007}
\received[revised]{12 March 2009}
\received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
ACM's consolidated article template, introduced in 2017, provides a
consistent \LaTeX\ style for use across ACM publications, and
incorporates accessibility and metadata-extraction functionality
necessary for future Digital Library endeavors. Numerous ACM and
SIG-specific \LaTeX\ templates have been examined, and their unique
features incorporated into this single new template.

If you are new to publishing with ACM, this document is a valuable
guide to the process of preparing your work for publication. If you
have published with ACM before, this document provides insight and
instruction into more recent changes to the article template.

The ``\verb|acmart|'' document class can be used to prepare articles
for any ACM publication --- conference or journal, and for any stage
of publication, from review to final ``camera-ready'' copy, to the
author's own version, with {\itshape very} few changes to the source.

\section{Template Overview}
As noted in the introduction, the ``\verb|acmart|'' document class can
be used to prepare many different kinds of documentation --- a
double-anonymous initial submission of a full-length technical paper, a
two-page SIGGRAPH Emerging Technologies abstract, a ``camera-ready''
journal article, a SIGCHI Extended Abstract, and more --- all by
selecting the appropriate {\itshape template style} and {\itshape
  template parameters}.

This document will explain the major features of the document
class. For further information, the {\itshape \LaTeX\ User's Guide} is
available from
\url{https://www.acm.org/publications/proceedings-template}.

\subsection{Template Styles}

The primary parameter given to the ``\verb|acmart|'' document class is
the {\itshape template style} which corresponds to the kind of publication
or SIG publishing the work. This parameter is enclosed in square
brackets and is a part of the {\verb|documentclass|} command:
\begin{verbatim}
  \documentclass[STYLE]{acmart}
\end{verbatim}

Journals use one of three template styles. All but three ACM journals
use the {\verb|acmsmall|} template style:
\begin{itemize}
\item {\texttt{acmsmall}}: The default journal template style.
\item {\texttt{acmlarge}}: Used by JOCCH and TAP.
\item {\texttt{acmtog}}: Used by TOG.
\end{itemize}

The majority of conference proceedings documentation will use the {\verb|acmconf|} template style.
\begin{itemize}
\item {\texttt{sigconf}}: The default proceedings template style.
\item{\texttt{sigchi}}: Used for SIGCHI conference articles.
\item{\texttt{sigplan}}: Used for SIGPLAN conference articles.
\end{itemize}

\subsection{Template Parameters}

In addition to specifying the {\itshape template style} to be used in
formatting your work, there are a number of {\itshape template parameters}
which modify some part of the applied template style. A complete list
of these parameters can be found in the {\itshape \LaTeX\ User's Guide.}

Frequently-used parameters, or combinations of parameters, include:
\begin{itemize}
\item {\texttt{anonymous,review}}: Suitable for a ``double-anonymous''
  conference submission. Anonymizes the work and includes line
  numbers. Use with the \texttt{\acmSubmissionID} command to print the
  submission's unique ID on each page of the work.
\item{\texttt{authorversion}}: Produces a version of the work suitable
  for posting by the author.
\item{\texttt{screen}}: Produces colored hyperlinks.
\end{itemize}

This document uses the following string as the first command in the
source file:
\begin{verbatim}
\documentclass[sigconf]{acmart}
\end{verbatim}

\section{Modifications}

Modifying the template --- including but not limited to: adjusting
margins, typeface sizes, line spacing, paragraph and list definitions,
and the use of the \verb|\vspace| command to manually adjust the
vertical spacing between elements of your work --- is not allowed.

{\bfseries Your document will be returned to you for revision if
  modifications are discovered.}

\section{Typefaces}

The ``\verb|acmart|'' document class requires the use of the
``Libertine'' typeface family. Your \TeX\ installation should include
this set of packages. Please do not substitute other typefaces. The
``\verb|lmodern|'' and ``\verb|ltimes|'' packages should not be used,
as they will override the built-in typeface families.

\section{Title Information}

The title of your work should use capital letters appropriately -
\url{https://capitalizemytitle.com/} has useful rules for
capitalization. Use the {\verb|title|} command to define the title of
your work. If your work has a subtitle, define it with the
{\verb|subtitle|} command.  Do not insert line breaks in your title.

If your title is lengthy, you must define a short version to be used
in the page headers, to prevent overlapping text. The \verb|title|
command has a ``short title'' parameter:
\begin{verbatim}
  \title[short title]{full title}
\end{verbatim}

\section{Authors and Affiliations}

Each author must be defined separately for accurate metadata
identification.  As an exception, multiple authors may share one
affiliation. Authors' names should not be abbreviated; use full first
names wherever possible. Include authors' e-mail addresses whenever
possible.

Grouping authors' names or e-mail addresses, or providing an ``e-mail
alias,'' as shown below, is not acceptable:
\begin{verbatim}
  \author{Brooke Aster, David Mehldau}
  \email{dave,judy,steve@university.edu}
  \email{firstname.lastname@phillips.org}
\end{verbatim}

The \verb|authornote| and \verb|authornotemark| commands allow a note
to apply to multiple authors --- for example, if the first two authors
of an article contributed equally to the work.

If your author list is lengthy, you must define a shortened version of
the list of authors to be used in the page headers, to prevent
overlapping text. The following command should be placed just after
the last \verb|\author{}| definition:
\begin{verbatim}
  \renewcommand{\shortauthors}{McCartney, et al.}
\end{verbatim}
Omitting this command will force the use of a concatenated list of all
of the authors' names, which may result in overlapping text in the
page headers.

The article template's documentation, available at
\url{https://www.acm.org/publications/proceedings-template}, has a
complete explanation of these commands and tips for their effective
use.

Note that authors' addresses are mandatory for journal articles.

\section{Rights Information}

Authors of any work published by ACM will need to complete a rights
form. Depending on the kind of work, and the rights management choice
made by the author, this may be copyright transfer, permission,
license, or an OA (open access) agreement.

Regardless of the rights management choice, the author will receive a
copy of the completed rights form once it has been submitted. This
form contains \LaTeX\ commands that must be copied into the source
document. When the document source is compiled, these commands and
their parameters add formatted text to several areas of the final
document:
\begin{itemize}
\item the ``ACM Reference Format'' text on the first page.
\item the ``rights management'' text on the first page.
\item the conference information in the page header(s).
\end{itemize}

Rights information is unique to the work; if you are preparing several
works for an event, make sure to use the correct set of commands with
each of the works.

The ACM Reference Format text is required for all articles over one
page in length, and is optional for one-page articles (abstracts).

\section{CCS Concepts and User-Defined Keywords}

Two elements of the ``acmart'' document class provide powerful
taxonomic tools for you to help readers find your work in an online
search.

The ACM Computing Classification System ---
\url{https://www.acm.org/publications/class-2012} --- is a set of
classifiers and concepts that describe the computing
discipline. Authors can select entries from this classification
system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the
commands to be included in the \LaTeX\ source.

User-defined keywords are a comma-separated list of words and phrases
of the authors' choosing, providing a more flexible way of describing
the research being presented.

CCS concepts and user-defined keywords are required for for all
articles over two pages in length, and are optional for one- and
two-page articles (or abstracts).

\section{Sectioning Commands}

Your work should use standard \LaTeX\ sectioning commands:
\verb|section|, \verb|subsection|, \verb|subsubsection|, and
\verb|paragraph|. They should be numbered; do not remove the numbering
from the commands.

Simulating a sectioning command by setting the first word or words of
a paragraph in boldface or italicized text is {\bfseries not allowed.}

\section{Tables}

The ``\verb|acmart|'' document class includes the ``\verb|booktabs|''
package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
high-quality tables.

Table captions are placed {\itshape above} the table.

Because tables cannot be split across pages, the best placement for
them is typically the top of the page nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and the
table caption.  The contents of the table itself must go in the
\textbf{tabular} environment, to be aligned properly in rows and
columns, with the desired horizontal and vertical rules.  Again,
detailed instructions on \textbf{tabular} material are found in the
\textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table~\ref{tab:freq} is included in the input file; compare the
placement of the table here with the table in the printed output of
this document.

\begin{table}
  \caption{Frequency of Special Characters}
  \label{tab:freq}
  \begin{tabular}{ccl}
    \toprule
    Non-English or Math&Frequency&Comments\\
    \midrule
    \O & 1 in 1,000& For Swedish names\\
    $\pi$ & 1 in 5& Common in math\\
    \$ & 4 in 5 & Used in business\\
    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
  \bottomrule
\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of the page's
live area, use the environment \textbf{table*} to enclose the table's
contents and the table caption.  As with a single-column table, this
wide table will ``float'' to a location deemed more
desirable. Immediately following this sentence is the point at which
Table~\ref{tab:commands} is included in the input file; again, it is
instructive to compare the placement of the table here with the table
in the printed output of this document.

\begin{table*}
  \caption{Some Typical Commands}
  \label{tab:commands}
  \begin{tabular}{ccl}
    \toprule
    Command &A Number & Comments\\
    \midrule
    \texttt{{\char'134}author} & 100& Author \\
    \texttt{{\char'134}table}& 300 & For tables\\
    \texttt{{\char'134}table*}& 400& For wider tables\\
    \bottomrule
  \end{tabular}
\end{table*}

Always use midrule to separate table header rows from data rows, and
use it only for this purpose. This enables assistive technologies to
recognise table headers and support their users in navigating tables
more easily.

\section{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of the three are
discussed in the next sections.

\subsection{Inline (In-text) Equations}
A formula that appears in the running text is called an inline or
in-text formula.  It is produced by the \textbf{math} environment,
which can be invoked with the usual
\texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
and structures, from $\alpha$ to $\omega$, available in
\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
examples of in-text equations in context. Notice how this equation:
\begin{math}
  \lim_{n\rightarrow \infty}x=0
\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsection{Display Equations}
A numbered display equation---one set off by vertical space from the
text and centered horizontally---is produced by the \textbf{equation}
environment. An unnumbered display equation is produced by the
\textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols and
structures available in \LaTeX\@; this section will just give a couple
of examples of display equations in context.  First, consider the
equation, shown as an inline equation above:
\begin{equation}
  \lim_{n\rightarrow \infty}x=0
\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}
  \sum_{i=0}^{\infty} x + 1
\end{displaymath}
and follow it with another numbered equation:
\begin{equation}
  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\section{Figures}

The ``\verb|figure|'' environment should be used for figures. One or
more images can be placed within a figure. If your figure contains
third-party material, you must clearly identify it as such, as shown
in the example below.
\begin{figure}[h]
  \centering
  %\includegraphics[width=\linewidth]{sample-franklin}
  \caption{1907 Franklin Model D roadster. Photograph by Harris \&
    Ewing, Inc. [Public domain], via Wikimedia
    Commons. (\url{https://goo.gl/VLCRBB}).}
  \Description{A woman and a girl in white dresses sit in an open car.}
\end{figure}

Your figures should contain a caption which describes the figure to
the reader.

Figure captions are placed {\itshape below} the figure.

Every figure should also have a figure description unless it is purely
decorative. These descriptions convey what’s in the image to someone
who cannot see it. They are also used by search engine crawlers for
indexing images, and when images cannot be loaded.

A figure description must be unformatted plain text less than 2000
characters long (including spaces).  {\bfseries Figure descriptions
  should not repeat the figure caption – their purpose is to capture
  important information that is not already provided in the caption or
  the main text of the paper.} For figures that convey important and
complex new information, a short text description may not be
adequate. More complex alternative descriptions can be placed in an
appendix and referenced in a short figure description. For example,
provide a data table capturing the information in a bar chart, or a
structured list representing a graph.  For additional information
regarding how best to write figure descriptions and why doing this is
so important, please see
\url{https://www.acm.org/publications/taps/describing-figures/}.

\subsection{The ``Teaser Figure''}

A ``teaser figure'' is an image, or set of images in one figure, that
are placed after all author and affiliation information, and before
the body of the article, spanning the page. If you wish to have such a
figure in your article, place the command immediately before the
\verb|\maketitle| command:
\begin{verbatim}
  \begin{teaserfigure}
    \includegraphics[width=\textwidth]{sampleteaser}
    \caption{figure caption}
    \Description{figure description}
  \end{teaserfigure}
\end{verbatim}

\section{Citations and Bibliographies}

The use of \BibTeX\ for the preparation and formatting of one's
references is strongly recommended. Authors' names should be complete
--- use full first names (``Donald E. Knuth'') not initials
(``D. E. Knuth'') --- and the salient identifying features of a
reference should be included: title, year, volume, number, pages,
article DOI, etc.

The bibliography is included in your source document with these two
commands, placed just before the \verb|\end{document}| command:
\begin{verbatim}
  \bibliographystyle{ACM-Reference-Format}
  \bibliography{bibfile}
\end{verbatim}
where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
suffix, of the \BibTeX\ file.

Citations and references are numbered by default. A small number of
ACM publications have citations and references formatted in the
``author year'' style; for these exceptions, please include this
command in the {\bfseries preamble} (before the command
``\verb|\begin{document}|'') of your \LaTeX\ source:
\begin{verbatim}
  \citestyle{acmauthoryear}
\end{verbatim}

  Some examples.  A paginated journal article \cite{Abril07}, an
  enumerated journal article \cite{Cohen07}, a reference to an entire
  issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
  monograph/whole book in a series (see 2a in spec. document)
  \cite{Harel79}, a divisible-book such as an anthology or compilation
  \cite{Editor00} followed by the same example, however we only output
  the series if the volume number is given \cite{Editor00a} (so
  Editor00a's series should NOT be present since it has no vol. no.),
  a chapter in a divisible book \cite{Spector90}, a chapter in a
  divisible book in a series \cite{Douglass98}, a multi-volume work as
  book \cite{Knuth97}, a couple of articles in a proceedings (of a
  conference, symposium, workshop for example) (paginated proceedings
  article) \cite{Andler79, Hagerup1993}, a proceedings article with
  all possible elements \cite{Smith10}, an example of an enumerated
  proceedings article \cite{VanGundy07}, an informally published work
  \cite{Harel78}, a couple of preprints \cite{Bornmann2019,
    AnzarootPBM14}, a doctoral dissertation \cite{Clarkson85}, a
  master's thesis: \cite{anisi03}, an online document / world wide web
  resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game
  (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05}
  and (Case 3) a patent \cite{JoeScientist001}, work accepted for
  publication \cite{rous08}, 'YYYYb'-test for prolific author
  \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
  contain 'duplicate' DOI and URLs (some SIAM articles)
  \cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
  multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
  couple of citations with DOIs:
  \cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
  citations: \cite{TUGInstmem, Thornburg01, CTANacmart}.
  Artifacts: \cite{R} and \cite{UMassCitations}.

\section{Acknowledgments}

Identification of funding sources and other support, and thanks to
individuals and groups that assisted in the research and the
preparation of the work should be included in an acknowledgment
section, which is placed just before the reference section in your
document.

This section has a special environment:
\begin{verbatim}
  \begin{acks}
  ...
  \end{acks}
\end{verbatim}
so that the information contained therein can be more easily collected
during the article metadata extraction phase, and to ensure
consistency in the spelling of the section heading.

Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.

\section{Appendices}

If your work needs an appendix, add it before the
``\verb|\end{document}|'' command at the conclusion of your source
document.

Start the appendix with the ``\verb|appendix|'' command:
\begin{verbatim}
  \appendix
\end{verbatim}
and note that in the appendix, sections are lettered, not
numbered. This document has two appendices, demonstrating the section
and subsection identification method.

\section{Multi-language papers}

Papers may be written in languages other than English or include
titles, subtitles, keywords and abstracts in different languages (as a
rule, a paper in a language other than English should include an
English title and an English abstract).  Use \verb|language=...| for
every language used in the paper.  The last language indicated is the
main language of the paper.  For example, a French paper with
additional titles and abstracts in English and German may start with
the following command
\begin{verbatim}
\documentclass[sigconf, language=english, language=german,
               language=french]{acmart}
\end{verbatim}

The title, subtitle, keywords and abstract will be typeset in the main
language of the paper.  The commands \verb|\translatedXXX|, \verb|XXX|
begin title, subtitle and keywords, can be used to set these elements
in the other languages.  The environment \verb|translatedabstract| is
used to set the translation of the abstract.  These commands and
environment have a mandatory first argument: the language of the
second argument.  See \verb|sample-sigconf-i13n.tex| file for examples
of their usage.

\section{SIGCHI Extended Abstracts}

The ``\verb|sigchi-a|'' template style (available only in \LaTeX\ and
not in Word) produces a landscape-orientation formatted article, with
a wide left margin. Three environments are available for use with the
``\verb|sigchi-a|'' template style, and produce formatted output in
the margin:
\begin{description}
\item[\texttt{sidebar}:]  Place formatted text in the margin.
\item[\texttt{marginfigure}:] Place a figure in the margin.
\item[\texttt{margintable}:] Place a table in the margin.
\end{description}

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
To Robert, for the bagels and explaining CMYK and color spaces.
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\section{Research Methods}

\subsection{Part One}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
lacinia dolor. Integer ultricies commodo sem nec semper.

\subsection{Part Two}

Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
eros. Vivamus non purus placerat, scelerisque diam eu, cursus
ante. Etiam aliquam tortor auctor efficitur mattis.

\section{Online Resources}

Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
enim maximus. Vestibulum gravida massa ut felis suscipit
congue. Quisque mattis elit a risus ultrices commodo venenatis eget
dui. Etiam sagittis eleifend elementum.

Nam interdum magna at lectus dignissim, ac dignissim lorem
rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
massa et mattis lacinia.

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.

### 2024-11-30 paper v2/paper v2.tex
feat: change paper format & add data analysis section

%%%%%%%% ICML 2021 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%

\documentclass{article}

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables
\usepackage{hyperref}
% the following package enables the use of multiline comments:
\usepackage{comment}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
%\usepackage{icml2021}

% If accepted, instead use the following line for the camera-ready submission:
\usepackage[accepted]{icml2021}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Board games group recomendation using metadata}

\begin{document}

\twocolumn[
\icmltitle{Board games group recomendation using metadata}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2021
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Eduardo Salinas}{to}
\icmlauthor{Alfonso Badilla}{to}
\icmlauthor{Nicolás Gutiérrez}{to}
\end{icmlauthorlist}

\icmlaffiliation{to}{Department of Computation, Pontificia Universidad Católica de Chile, Santiago, Chile}

\icmlcorrespondingauthor{Eduardo Salinas}{esalinasbarros@uc.cl}
\icmlcorrespondingauthor{Alfonso Badilla}{alfonso.badilla@uc.cl}
\icmlcorrespondingauthor{Nicolás Gutiérrez}{njgutierrez@uc.cl}

% You may provide any keywords that you find helpful for describing your paper; these are used to populate the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{Machine Learning, ICML, Group recommendation, Metadata}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

\printAffiliationsAndNotice{} % otherwise use the standard text.

\begin{abstract}
In this document we propose a group recommendation system for board games using data from the Board Game Geek website, which includes attributes such as the number of players, average playtime, complexity, and user ratings.
% todo: agregar explicación groso modo del algoritmo
The implementation employs Python and the scikit-learn library, offering an efficient and scalable solution for group-based game recommendations.
\end{abstract}

\section{Motivation and state of the art}
\label{Motivation}

Board games have been a popular form of entertainment for centuries, and their popularity has only increased in recent years. Even as digital games have become more prevalent, board games have remained a favorite pastime for many people, and the market for board games has grown significantly giving rise to plataforms such as Board Game Geek (BGG). This website is a popular platform for board game enthusiasts, where users can rate and review games, as well as access information about them.

This plataform has a vast amount of data that has been used in the past to create recommendation systems. For example, 
%\cite{DBLP:journals/corr/abs-2006-07053} 
proposed a recommendation system for board games based on user ratings, using a collaborative filtering approach. However, this approach does not take into account the characteristics of many of the games, that are supposed to be played in groups. As such, we propose a group recommendation system to make recommendations tailored to groups of players.

Another example of group recommendation systems is the work of %\cite{DBLP:journals/corr/abs-2006-07053}
, who proposed a group recommendation system for tourism and movies\footnote{This work is hosted on a github repo and is intended to be used as a basic tutorial.}.
This system uses regular clustering algorithms to group users based on their ratings, such as user k-nearest, item k-nearest and support vector decomposition, and then recommends movies and touristic locations based on the preferences of the group.

\subsection{Proposal}

We propose a refined version of that last system, using metadata-centric clustering to group board games based on their characteristics, and then recommending games to groups of players based on their summed preferences used the functions detailed on the previous work. This system is designed to be more accuarate than the previous one, as it takes into account the characteristics of the games, and not just the ratings of the users.

\section{Data analysis}

The dataset that will be used for this project has 10 different files, each one tith different information about board games. The main file is \texttt{user\_ratings.csv}, which contains 19 million rows. Due to the limited processing power of Google Collab, we will use only 50,000 rows for the initial training. The other files are:

\begin{itemize}
    \item \texttt{games.csv}: This file contains 22 attributes about the rated games and includes 22,000 different games. Among the attributes, we have aspects such as the game description, which could be evaluated using language models, the year of publication (which distributes as shown in figure \ref{fig:publishYear}), the average rating received, among others.
    \item \texttt{ratings\_distribution.csv}: Contains the total ratings for each board game. In general, these ratings distribute as shown in figure \ref{fig:distribucionRatings}.
    \item \texttt{temes.csv}: Contains the themes of each game as a binary flag. The 50 most common themes can be seen in the graph in figure \ref{fig:tematicasComunes}.
    \item \texttt{mechanics.csv}: Contains the game mechanics as a binary flag. The 50 most common mechanics can be seen in the graph in figure \ref{fig:mecanicasComunes}.
    \item \texttt{subcategories.csv}: Contains the subcategories of each game in binary flag format.
    \item \texttt{artists\_reduced.csv}: Contains information about which artist created a certain game in binary flag format. Only artists with more than 3 games created are considered.
    \item \texttt{designer\_reduced.csv}: Contains information about which designer designed a certain game in binary flag format. Only designers with more than 3 games created are considered.
    \item \texttt{publishers\_reduced.csv}: Contains information about the companies that sell these games in binary flag format. Only companies that sell more than 3 games are considered.

\end{itemize}

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{publishYear.png}
    \caption{Juegos por año de publicación}
    \label{fig:publishYear}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\linewidth]{distribucionRatings.png}
    \caption{Distribución de ratings de juegos}
    \label{fig:distribucionRatings}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\linewidth]{tematicasComunes.png}
    \caption{Temáticas más comunes en el dataset}
    \label{fig:tematicasComunes}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\linewidth]{mecanicasComunes.png}
    \caption{Mecánicas más comunes en el dataset}
    \label{fig:mecanicasComunes}
\end{figure}

% You can use footnotes\footnote{Footnotes should be complete sentences.} to provide readers with additional information about a topic without interrupting the flow of the paper.

% \begin{figure}[ht]
% \vskip 0.2in
% \begin{center}
% \centerline{\includegraphics[width=\columnwidth]{icml_numpapers}}
% \caption{Historical locations and number of accepted papers for International
% Machine Learning Conferences (ICML 1993 -- ICML 2008) and International
% Workshops on Machine Learning (ML 1988 -- ML 1992). At the time this figure was
% produced, the number of accepted papers for ICML 2008 was unknown and instead
% estimated.}
% \label{icml-historical}
% \end{center}
% \vskip -0.2in
% \end{figure}

\subsection{Figures}

You may want to include figures in the paper to illustrate your approach and results. Such artwork should be centered, legible, and separated from the text. Lines should be dark and at least 0.5~points thick for purposes of reproduction, and text should not appear on a gray background.

Label all distinct components of each figure. If the figure takes the form of a graph, then give a name for each axis and include a legend that briefly describes each curve. Do not include a title inside the figure; instead, the caption should serve this function.

%Number figures sequentially, placing the figure number and caption \emph{after} the graphics, with at least 0.1~inches of space before

%You may float figures to the top or bottom of a column, and you may set wide figures across both columns (use the environment \texttt{figure*} in \LaTeX). Always place two-column figures at the top or bottom of the page.

% Ejemplo de algoritmo (creo que no usaremos esto)
% \begin{algorithm}[tb]
%    \caption{Bubble Sort}
%    \label{alg:example}
% \begin{algorithmic}
%    \STATE {\bfseries Input:} data $x_i$, size $m$
%    \REPEAT
%    \STATE Initialize $noChange = true$.
%    \FOR{$i=1$ {\bfseries to} $m-1$}
%    \IF{$x_i > x_{i+1}$}
%    \STATE Swap $x_i$ and $x_{i+1}$
%    \STATE $noChange = false$
%    \ENDIF
%    \ENDFOR
%    \UNTIL{$noChange$ is $true$}
% \end{algorithmic}
% \end{algorithm}

% Ejemplo de tabla
% \begin{table}[t]
% \caption{Classification accuracies for naive Bayes and flexible
% Bayes on various data sets.}
% \label{sample-table}
% \vskip 0.15in
% \begin{center}
% \begin{small}
% \begin{sc}
% \begin{tabular}{lcccr}
% \toprule
% Data set & Naive & Flexible & Better? \\
% \midrule
% Breast    & 95.9$\pm$ 0.2& 96.7$\pm$ 0.2& $\surd$ \\
% Cleveland & 83.3$\pm$ 0.6& 80.0$\pm$ 0.6& $\times$\\
% Glass2    & 61.9$\pm$ 1.4& 83.8$\pm$ 0.7& $\surd$ \\
% Credit    & 74.8$\pm$ 0.5& 78.3$\pm$ 0.6&         \\
% Horse     & 73.3$\pm$ 0.9& 69.7$\pm$ 1.0& $\times$\\
% Meta      & 67.1$\pm$ 0.6& 76.5$\pm$ 0.5& $\surd$ \\
% Pima      & 75.1$\pm$ 0.6& 73.9$\pm$ 0.5&         \\
% Vehicle   & 44.9$\pm$ 0.6& 61.5$\pm$ 0.4& $\surd$ \\
% \bottomrule
% \end{tabular}
% \end{sc}
% \end{small}
% \end{center}
% \vskip -0.1in
% \end{table}

\subsection{Citations and References}

Please use APA reference format regardless of your formatter
or word processor. If you rely on the \LaTeX\/ bibliographic
facility, use \texttt{natbib.sty} and \texttt{icml2021.bst}
included in the style-file package to obtain this format.

Citations within the text should include the authors' last names and
year. If the authors' names are included in the sentence, place only
the year in parentheses, for example when referencing Arthur Samuel's
pioneering work \yrcite{Samuel59}. Otherwise place the entire
reference in parentheses with the authors and year separated by a
comma \cite{Samuel59}. List multiple references separated by
semicolons \cite{kearns89,Samuel59,mitchell80}. Use the `et~al.'
construct only for citations with three or more authors or after
listing all authors to a publication in an earlier reference \cite{MachineLearningI}.

Authors should cite their own work in the third person
in the initial version of their paper submitted for blind review.
Please refer to Section~\ref{author info} for detailed instructions on how to
cite your own papers.

Use an unnumbered first-level section heading for the references, and use a
hanging indent style, with the first line of the reference flush against the
left margin and subsequent lines indented by 10 points. The references at the
end of this document give examples for journal articles \cite{Samuel59},
conference publications \cite{langley00}, book chapters \cite{Newell81}, books
\cite{DudaHart2nd}, edited volumes \cite{MachineLearningI}, technical reports
\cite{mitchell80}, and dissertations \cite{kearns89}.

Alphabetize references by the surnames of the first authors, with
single author entries preceding multiple author entries. Order
references for the same authors by year of publication, with the
earliest first. Make sure that each reference includes all relevant
information (e.g., page numbers).

Please put some effort into making references complete, presentable, and
consistent. If using bibtex, please protect capital letters of names and
abbreviations in titles, for example, use \{B\}ayesian or \{L\}ipschitz
in your .bib file.

% In the unusual situation where you want a paper to appear in the
% references without citing it in the main text, use \nocite
\nocite{langley00}

\bibliography{example_paper}
\bibliographystyle{icml2021}

\end{document}

### 2024-11-30 paper v2/paper v2.tex
feat: defined sections of the paper

@@
 
+\section{Methodology}
+
+The methodology for this project is divided into three main stages: data preprocessing, group making, and recommendation.
+The first stage consists of reducing the data and selecting the most relevant attributes.
+The second stage consists of grouping the games based on their characteristics, and the third stage consists of recommending games to groups of players based on their preferences.
+
+\subsection{Data preprocessing}
+
+\subsection{Group making}
+
+\subsection{Recommendation}
+
+\section{Sensitivity Analysis}
+
+\section{Results}
+
+\section{Conclusions}
+
 % You can use footnotes\footnote{Footnotes should be complete sentences.} to provide readers with additional information about a topic without interrupting the flow of the paper.
@@
 
-\subsection{Figures}
+\section{Figures}
 

### 2024-11-30 metrics.py
Evaluation of group recommendation

@@ -1,0 +1,121 @@
+import ast
+import numpy as np
+import pandas as pd
+
+def parse_group_data(groups_df):
+    parsed_data = []
+    
+    for _, row in groups_df.head(250).iterrows():
+        group_members = ast.literal_eval(row['group_members'])
+        recommendations = ast.literal_eval(row['recommendations'])
+        parsed_data.append((group_members, recommendations))
+    
+    print(f"\nTotal groups parsed: {len(parsed_data)}")
+    return parsed_data
+
+def calculate_relevance(train_df, user_id, item_id):
+    rating = train_df[
+        (train_df['Username'] == user_id) & 
+        (train_df['BGGId'] == item_id)
+    ]['Rating']
+    
+    return rating.iloc[0] if not rating.empty else 0
+
+def calculate_group_metrics(parsed_data, train_df, k=5, threshold=3.5):
+    total_precision = 0
+    total_recall = 0
+    total_ndcg = 0
+    total_relevance = 0
+    total_groups = len(parsed_data)
+    total_hits = 0
+    total_possible_hits = 0
+    all_group_sizes = []
+    
+    for i, (group_members, recommendations) in enumerate(parsed_data):
+        print(f"\rProcessing Group {i+1}/{total_groups}", end="")
+        recommended_items = recommendations[:k]
+        group_relevance = []
+        group_hits = 0
+        for item_id in recommended_items:
+            member_ratings = []
+            for user_id in group_members:
+                rating = calculate_relevance(train_df, user_id, item_id)
+                if rating > 0:
+                    member_ratings.append(rating)
+                    group_hits += 1
+                total_possible_hits += 1
+            avg_rating = np.mean(member_ratings) if member_ratings else 0
+            group_relevance.append(avg_rating)
+        
+        total_hits += group_hits
+        
+        binary_relevance = [1 if rel >= threshold else 0 for rel in group_relevance]
+        
+        total_relevant = 0
+        unique_items = train_df['BGGId'].unique()
+        for item_id in unique_items:
+            member_ratings = []
+            for user_id in group_members:
+                rating = calculate_relevance(train_df, user_id, item_id)
+                if rating > 0:
+                    member_ratings.append(rating)
[…]

### 2024-12-02 metrics.py
Evaluation of group recommendation

@@ -5,7 +5,7 @@ import pandas as pd
 def parse_group_data(groups_df):
     parsed_data = []
     
-    for _, row in groups_df.head(250).iterrows():
+    for _, row in groups_df.head(300).iterrows():
         group_members = ast.literal_eval(row['group_members'])
         recommendations = ast.literal_eval(row['recommendations'])
         parsed_data.append((group_members, recommendations))
@@ -21,19 +21,28 @@ def calculate_relevance(train_df, user_id, item_id):
     
     return rating.iloc[0] if not rating.empty else 0
 
-def calculate_group_metrics(parsed_data, train_df, k=5, threshold=3.5):
+def calculate_group_metrics(parsed_data, train_df, k=5, threshold=5):
     total_precision = 0
     total_recall = 0
     total_ndcg = 0
     total_relevance = 0
+    total_diversity = 0
     total_groups = len(parsed_data)
     total_hits = 0
     total_possible_hits = 0
-    all_group_sizes = []
+    
+    # Keep track of all recommended items
+    all_recommended_items = set()
+    total_recommendations = 0
     
     for i, (group_members, recommendations) in enumerate(parsed_data):
         print(f"\rProcessing Group {i+1}/{total_groups}", end="")
         recommended_items = recommendations[:k]
+        
+        # Add to global set of recommended items
+        all_recommended_items.update(recommended_items)
+        total_recommendations += len(recommended_items)
+        
         group_relevance = []
         group_hits = 0
         for item_id in recommended_items:
@@ -78,16 +87,15 @@ def calculate_group_metrics(parsed_data, train_df, k=5, threshold=3.5):
         total_ndcg += ndcg
         total_relevance += np.mean(group_relevance) if group_relevance else 0
     
+    # Calculate global diversity
+    total_diversity = len(all_recommended_items) / total_recommendations
+    
     avg_metrics = {
         "average_precision": total_precision / total_groups,
         "average_recall": total_recall / total_groups,
         "average_ndcg": total_ndcg / total_groups,
         "average_relevance": total_relevance / total_groups,
-        # "hit_rate": total_hits / total_possible_hits,
-        # "average_group_size": np.mean(all_group_sizes),
-        # "total_groups": total_groups,
[…]

### 2024-12-02 metrics.py
Evaluation of group recommendation

@@ -10,7 +10,7 @@ def parse_group_data(groups_df):
         recommendations = ast.literal_eval(row['recommendations'])
         parsed_data.append((group_members, recommendations))
     
-    print(f"\nTotal groups parsed: {len(parsed_data)}")
+    print(f"\nTotal: {len(parsed_data)}")
     return parsed_data
 
 def calculate_relevance(train_df, user_id, item_id):
@@ -31,15 +31,13 @@ def calculate_group_metrics(parsed_data, train_df, k=5, threshold=5):
     total_hits = 0
     total_possible_hits = 0
     
-    # Keep track of all recommended items
     all_recommended_items = set()
     total_recommendations = 0
     
     for i, (group_members, recommendations) in enumerate(parsed_data):
-        print(f"\rProcessing Group {i+1}/{total_groups}", end="")
+        print(f"\rProcesando {i+1}/{total_groups}", end="")
         recommended_items = recommendations[:k]
         
-        # Add to global set of recommended items
         all_recommended_items.update(recommended_items)
         total_recommendations += len(recommended_items)
         
@@ -87,7 +85,6 @@ def calculate_group_metrics(parsed_data, train_df, k=5, threshold=5):
         total_ndcg += ndcg
         total_relevance += np.mean(group_relevance) if group_relevance else 0
     
-    # Calculate global diversity
     total_diversity = len(all_recommended_items) / total_recommendations
     
     avg_metrics = {
@@ -95,7 +92,7 @@ def calculate_group_metrics(parsed_data, train_df, k=5, threshold=5):
         "average_recall": total_recall / total_groups,
         "average_ndcg": total_ndcg / total_groups,
         "average_relevance": total_relevance / total_groups,
-        "average_diversity": total_diversity,  # This is now global diversity
+        "average_diversity": total_diversity,  
     }
     
     return avg_metrics

### 2024-12-04 metrics.py
New metrics eval

@@ -98,9 +98,25 @@ def calculate_group_metrics(parsed_data, train_df, k=5, threshold=5):
     return avg_metrics
 
 train_df = pd.read_csv('train.csv')
-groups_df = pd.read_csv('RecSysProyectoFinal/group_recommendations.csv')
+# df = pd.read_csv('RecSysProyectoFinal/random_group_recommendations.csv', sep=';')
+# df.to_csv('RecSysProyectoFinal/random_group_recommendations.csv', index=False, quoting=1)
+groups_df = pd.read_csv('RecSysProyectoFinal/random_group_recommendations.csv')
 parsed_data = parse_group_data(groups_df)
 metrics = calculate_group_metrics(parsed_data, train_df, k=5)
+print("Random LightFM")
+print(f"Precision: {metrics['average_precision']:.6f}")
+print(f"Recall: {metrics['average_recall']:.6f}")
+print(f"nDCG: {metrics['average_ndcg']:.6f}")
+print(f"Relevance Score: {metrics['average_relevance']:.2f}")
+print(f"Diversity: {metrics['average_diversity']:.2f}")
+
+
+# df = pd.read_csv('RecSysProyectoFinal/similar_group_recommendations.csv', sep=';')
+# df.to_csv('RecSysProyectoFinal/similar_group_recommendations.csv', index=False, quoting=1)
+groups_df = pd.read_csv('RecSysProyectoFinal/similar_group_recommendations.csv')
+parsed_data = parse_group_data(groups_df)
+metrics = calculate_group_metrics(parsed_data, train_df, k=5)
+print("Similar LightFM")
 print(f"Precision: {metrics['average_precision']:.6f}")
 print(f"Recall: {metrics['average_recall']:.6f}")
 print(f"nDCG: {metrics['average_ndcg']:.6f}")
@@ -108,9 +124,24 @@ print(f"Relevance Score: {metrics['average_relevance']:.2f}")
 print(f"Diversity: {metrics['average_diversity']:.2f}")
 
 train_df = pd.read_csv('train.csv')
-groups_df = pd.read_csv('RecSysProyectoFinal/svd_group_recommendations.csv')
+# df = pd.read_csv('RecSysProyectoFinal/svd_random_group_recommendations.csv', sep=';')
+# df.to_csv('RecSysProyectoFinal/svd_random_group_recommendations.csv', index=False, quoting=1)
[…]

### 2024-12-04 metrics.py
Metricas listos

@@ -34,6 +34,15 @@ def calculate_group_metrics(parsed_data, train_df, k=5, threshold=5):
     all_recommended_items = set()
     total_recommendations = 0
     
+    total_novelty = 0
+    item_popularity = {}
+    total_items = len(train_df['BGGId'].unique())
+    for item_id in train_df['BGGId'].unique():
+        item_popularity[item_id] = len(train_df[train_df['BGGId'] == item_id]) / len(train_df)
+    
+    total_fairness = 0
+    total_serendipity = 0
+    
     for i, (group_members, recommendations) in enumerate(parsed_data):
         print(f"\rProcesando {i+1}/{total_groups}", end="")
         recommended_items = recommendations[:k]
@@ -84,6 +93,41 @@ def calculate_group_metrics(parsed_data, train_df, k=5, threshold=5):
         total_recall += recall
         total_ndcg += ndcg
         total_relevance += np.mean(group_relevance) if group_relevance else 0
+        
+        
+        group_novelty = 0
+        for item_id in recommended_items:
+            pop = item_popularity.get(item_id, 0)
+            if pop > 0:
+                group_novelty += -np.log2(pop)
+        group_novelty /= len(recommended_items) if recommended_items else 1
+        total_novelty += group_novelty
+        
+        # Calculate fairness (variance of satisfaction among group members)
+        member_satisfaction = []
+        for user_id in group_members:
+            user_satisfaction = 0
+            for item_id in recommended_items:
+                rating = calculate_relevance(train_df, user_id, item_id)
+                user_satisfaction += rating
+            user_satisfaction /= len(recommended_items) if recommended_items else 1
+            member_satisfaction.append(user_satisfaction)
+        
+        # Lower variance means higher fairness
+        fairness = 1 / (1 + np.var(member_satisfaction)) if member_satisfaction else 0
+        total_fairness += fairness
+        
+        # Calculate serendipity (unexpected relevant recommendations)
+        serendipity = 0
+        for item_id in recommended_items:
+            # Item is considered serendipitous if it's both relevant and novel
[…]

### 2024-12-06 paper v2/paper v2.tex
feat: bases for methodology & results on the paper

--- 2024-11-30 paper v2/paper v2.tex
@@
 
+\section{Methodology}
+
+The methodology for this project is divided into three main stages: data preprocessing, group making, and recommendation.
+
+\subsection{Data preprocessing}
+The first stage consists of reducing the dataset by not picking items without interactions and then dividing the...
+
+% Mostrar gráfico de las clases de usuarios y clases de ítems, además de mostrar cuantos no valen la pena.
+
+\subsection{Group making}
+
+% Insertar referencia y citar apropiadamente el trabajo.
+The second stage consists of grouping the users based on their characteristics by using a correlation matrix.
+The function used is an adaptation of the one used in the "" work, which is based on the cosine similarity between the users.
+The function is as follows:
+
+% Arreglar función
+\begin{equation}
+    %\text{similarity}(u, v) = \frac{\sum_{i=1}^{n} u_i \cdot v_i}{\sqrt{\sum_{i=1}^{n} u_i^2} \cdot \sqrt{\sum_{i=1}^{n} v_i^2}}
+\end{equation}
+
+\subsection{Recommendation}
+The third stage consists of recommending games to groups of players based on their preferences.
+Since the recommendations are supposed to be made to groups, the preferences of the group are calculated using a variety of methods, that can be described as follows:
+\begin{enumerate}
+    \item Simple average: The preferences of the group are calculated as the simple average of the preferences of the individual players.
+    \item Least misery: The preferences of the group are calculated as the minimum of the preferences of the individual players.
+    \item ... % asegurarse que esto está al día con lo que presentemos al final del trabajo
+\end{enumerate}
+
+\section{Results}
+
+To evaluate the results of our model, we will compare it to most popular as a sanity check, to iknn applied to the same dataset and to SVD applied to the same dataset.
+% Agregar AGREE si logramos hacerlo funcionar ahora que tenemos menos datos.
+
+The metrics that will be used to evaluate the performance of each model are the following:
+
+\begin{itemize}
+    \item Precision at k: This metric measures the proportion of recommended items that are relevant to the user, where relevance is defined as the items that the user has interacted with.
+    \item Recall at k: This metric measures the proportion of relevant items that are recommended to the user.
+    \item F1 score: This metric is the harmonic mean of precision and recall, and is used to measure the overall performance of the model.
+\end{itemize}
+
+\section{Sensitivity Analysis}
+
+\section{Conclusions}
+
 % You can use footnotes\footnote{Footnotes should be complete sentences.} to provide readers with additional information about a topic without interrupting the flow of the paper.

### 2024-12-08 paper v2/paper v2.tex
fix: evaluación de SVD

@@
 \section{Motivation and state of the art}
-\label{Motivation}
 
@@
 
+\subsection{State of the art and related work}
+
 This plataform has a vast amount of data that has been used in the past to create recommendation systems. For example, 
-%\cite{DBLP:journals/corr/abs-2006-07053} 
-proposed a recommendation system for board games based on user ratings, using a collaborative filtering approach.
+github user richengo \cite{richengo2024} proposed a recommendation system for board games based on user ratings, using a collaborative filtering approach.
 However, this approach does not take into account the characteristics of many of the games, that are supposed to be played in groups.
@@
 The main file is \texttt{user\_ratings.csv}, which contains 19 million rows.
-Due to the limited processing power of Google Collab, we will use only 50,000 rows for the initial training.
+Due to the limited processing power of Google Collab and our own local machines, we will use a reduced set of items and users for this work.
 The other files are:
@@
     \item \texttt{subcategories.csv}: Contains the subcategories of each game in binary flag format.
-    \item \texttt{artists\_reduced.csv}: Contains information about which artist created a certain game in binary flag format. Only artists with more than 3 games created are considered.
+    \item \texttt{artists\_reduced.csv}: Contains information about which artist created a certain game in binary flag format. Only artists with more than 3 games created will be considered.
     \item \texttt{designer\_reduced.csv}: Contains information about which designer designed a certain game in binary flag format. Only designers with more than 3 games created are considered.

### 2024-12-09 paper v2/paper v2.tex
feat: referencias paper
chore: additional and corrected references

@@
 Board games have been a popular form of entertainment for centuries, and their popularity has only increased in recent years.
-Even as digital games have become more prevalent, board games have remained a favorite pastime for many people, and the market for board games has grown significantly giving rise to plataforms such as Board Game Geek (BGG).
+Even as digital games have become more prevalent, board games have remained a favorite pastime for many people, and the market for board games has grown significantly giving rise to plataforms such as Board Game Geek\footnote{\cite{boardgamegeek}} (BGG).
 This website is a popular platform for board game enthusiasts, where users can rate and review games, as well as access information about them.
@@
 This plataform has a vast amount of data that has been used in the past to create recommendation systems. For example,
-github user richengo \cite{richengo2024} proposed a recommendation system for board games based on user ratings, using a collaborative filtering approach.
+github user richengo \yrcite{richengo2024} proposed a recommendation system for board games based on user ratings, using a collaborative filtering approach.
 However, this approach does not take into account the characteristics of many of the games, that are supposed to be played in groups.
@@
 
-Another example of group recommendation systems is the work of %\cite{DBLP:journals/corr/abs-2006-07053}
-, who proposed a group recommendation system for tourism and movies\footnote{This work is hosted on a github repo and is intended to be used as a basic tutorial.}.
-This system uses regular clustering algorithms to group users based on their ratings, such as user k-nearest, item k-nearest and support vector decomposition, and then recommends movies and touristic locations based on the preferences of the group.
+Another example of group recommendation systems is the work of \cite{group_recommenders_repo}, who proposed a group recommendation system for tourism and movies\footnote{This work is hosted on a github repo and is intended to be used as a basic tutorial.}.
+This system uses regular clustering algorithms to group users based on their ratings, such as user k-nearest, item k-nearest and support vector decomposition (SVD), and then recommends movies and touristic locations based on the preferences of the group.
 
@@
 
-The dataset that will be used for this project has 10 different files, each one tith different information about board games.
+The dataset that will be used for this project\footnote{\cite{board_games_kaggle}}, obtained from the website kaggle, has 10 different files, each one tith different information about board games.
 The main file is \texttt{user\_ratings.csv}, which contains 19 million rows.
@@
 year. If the authors' names are included in the sentence, place only
-the year in parentheses, for example when referencing Arthur Samuel's
-pioneering work \yrcite{Samuel59}. Otherwise place the entire
-reference in parentheses with the authors and year separated by a
-comma \cite{Samuel59}. List multiple references separated by
-semicolons \cite{kearns89,Samuel59,mitchell80}. Use the `et~al.'
-construct only for citations with three or more authors or after
-listing all authors to a publication in an earlier reference \cite{MachineLearningI}.
-
-Authors should cite their own work in the third person
-in the initial version of their paper submitted for blind review.
-Please refer to Section~\ref{author info} for detailed instructions on how to
-cite your own papers.
-
-Use an unnumbered first-level section heading for the references, and use a
-hanging indent style, with the first line of the reference flush against the
-left margin and subsequent lines indented by 10 points. The references at the
-end of this document give examples for journal articles \cite{Samuel59},
-conference publications \cite{langley00}, book chapters \cite{Newell81}, books
-\cite{DudaHart2nd}, edited volumes \cite{MachineLearningI}, technical reports
-\cite{mitchell80}, and dissertations \cite{kearns89}.
+the year in parentheses using yrcite.
 
@@
 % references without citing it in the main text, use \nocite
-\nocite{langley00}
+% \nocite{}
 
-\bibliography{example_paper}
+\bibliography{main}
 \bibliographystyle{icml2021}

### 2024-12-09 paper v2/paper v2.tex
fix: translated figure labels

@@
     \includegraphics[width=0.8\linewidth]{publishYear.png}
-    \caption{Juegos por año de publicación}
+    \caption{Amount of games (y axis) by year of publication(x axis)}
     \label{fig:publishYear}
@@
     \includegraphics[width=0.8\linewidth]{distribucionRatings.png}
-    \caption{Distribución de ratings de juegos}
+    \caption{Games rating distribution (1-10)}
     \label{fig:distribucionRatings}
@@
     \includegraphics[width=0.8\linewidth]{tematicasComunes.png}
-    \caption{Temáticas más comunes en el dataset}
+    \caption{Most common game themes on the dataset}
     \label{fig:tematicasComunes}
@@
     \includegraphics[width=0.8\linewidth]{mecanicasComunes.png}
-    \caption{Mecánicas más comunes en el dataset}
+    \caption{Most common game mechanics on the dataset}
     \label{fig:mecanicasComunes}
@@
 \subsection{Data preprocessing}
-The first stage consists of reducing the dataset by not picking items without interactions and then dividing the...
+The first stage consists of reducing the dataset. This is important because the amount of data makes it take too long to calculate any meaningful metrics.
+
+Part of the reduction is already done on the dataset by default, since there are no games with less than 20 ratings.
+Since we are going to also reduce the amount of users, we will delete games further up to 40 ratings, and then pick from 4 different game classes.
 
@@
 
-\section{Figures}
+% \section{Figures}
 
-You may want to include figures in the paper to illustrate your approach and results.
+% You may want to include figures in the paper to illustrate your approach and results.
 Such artwork should be centered, legible, and separated from the text.
@@
 
-Label all distinct components of each figure.
+% Label all distinct components of each figure.
 If the figure takes the form of a graph, then give a name for each axis and include a legend that briefly describes each curve.
@@
 
-\subsection{Citations and References}
+% \subsection{Citations and References}
 
-Please use APA reference format regardless of your formatter
-or word processor. If you rely on the \LaTeX\/ bibliographic
-facility, use \texttt{natbib.sty} and \texttt{icml2021.bst}
-included in the style-file package to obtain this format.
+% Please use APA reference format regardless of your formatter
+% or word processor. If you rely on the \LaTeX\/ bibliographic
+% facility, use \texttt{natbib.sty} and \texttt{icml2021.bst}
+% included in the style-file package to obtain this format.
 
-Citations within the text should include the authors' last names and
-year. If the authors' names are included in the sentence, place only
-the year in parentheses using yrcite.
+% Citations within the text should include the authors' last names and
+% year. If the authors' names are included in the sentence, place only
+% the year in parentheses using yrcite.
 
-Alphabetize references by the surnames of the first authors, with
-single author entries preceding multiple author entries. Order
-references for the same authors by year of publication, with the
-earliest first. Make sure that each reference includes all relevant
-information (e.g., page numbers).
+% Alphabetize references by the surnames of the first authors, with
+% single author entries preceding multiple author entries. Order
+% references for the same authors by year of publication, with the
+% earliest first. Make sure that each reference includes all relevant
+% information (e.g., page numbers).
 
-Please put some effort into making references complete, presentable, and
-consistent. If using bibtex, please protect capital letters of names and
-abbreviations in titles, for example, use \{B\}ayesian or \{L\}ipschitz
-in your .bib file.
+% Please put some effort into making references complete, presentable, and
+% consistent. If using bibtex, please protect capital letters of names and
+% abbreviations in titles, for example, use \{B\}ayesian or \{L\}ipschitz
+% in your .bib file.
 

### 2024-12-09 paper v2/paper v2.tex
feat: paper data preprocessing

@@
 Part of the reduction is already done on the dataset by default, since there are no games with less than 20 ratings.
-Since we are going to also reduce the amount of users, we will delete games further up to 40 ratings, and then pick from 4 different game classes.
+Since we are going to also reduce the amount of users, we will delete games further up to 40 ratings, and then pick from 4 different game classes like the ones shown on figure \ref{fig:reviewsPorJuego}.
 
-% Mostrar gráfico de las clases de usuarios y clases de ítems, además de mostrar cuantos no valen la pena.
+With that in mind, we will reduce the amount of games to 1500, from the original $>$10k by taking 375 games from each class.
+
+We will also reduce the amount of users to 15k, from the original $>$300k, first deleting all users with only 1 review.
+However, the classes will be more complicated to build, since we will be deleting a lot of games, so just making classes like in figure \ref{fig:reviewsPorUsuario} will not be enough.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\linewidth]{ReviewsPorJuego.png}
+    \caption{Amount of reviews per game}
+    \label{fig:reviewsPorJuego}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\linewidth]{reviewsPorUsuario.png}
+    \caption{Amount of reviews per user (full dataset)}
+    \label{fig:reviewsPorUsuario}
+\end{figure}
+
+To do that, we will have to make the graph after selecting the games from each class.
+This graph can be seen in figure \ref{fig:reviewsPorUsuario2}, and as seen on that figure, the classes have shifted due to less reviews overall.
+We will then make 3 classes as shown, and select 5k users from each class.
+Note that we will not select users with less than 3 reviews.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\linewidth]{reviewsPorUsuario2.png}
+    \caption{Amount of reviews per user (1500 items)}
+    \label{fig:reviewsPorUsuario2}
+\end{figure}
+
+The final dataset will have 1500 games and 15k users, and will be used for the rest of the project.
+The dataset will be split into a training and a test set, with 80\% of the data being used for training and 20\% for testing.
+
+The final amount of reviews per user and per game can be seen in figure \ref{fig:reviewsPorUsuarioFinal} and figure \ref{fig:reviewsPorJuegoFinal} respectively.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\linewidth]{reviewsPorUsuarioFinal.png}
+    \caption{Amount of reviews per user (after full reduction)}
+    \label{fig:reviewsPorUsuarioFinal}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\linewidth]{ReviewsPorJuegoFinal.png}
+    \caption{Amount of reviews per game (after full reduction)}
+    \label{fig:reviewsPorJuegoFinal}
+\end{figure}
+
+With this, we expect to make a more efficient solution for group-based game recommendations, so that it can run with regular hardware.
 

### 2024-12-09 paper v2/paper v2.tex
fix: percentages of training and testing
feat: agregadas tablas de resultados

@@
 The final dataset will have 1500 games and 15k users, and will be used for the rest of the project.
-The dataset will be split into a training and a test set, with 80\% of the data being used for training and 20\% for testing.
+The dataset will be split into a training and a test set, with 75\% of the data being used for training and 25\% for testing.
 
@@
 \end{itemize}
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{|l|l|l|l|}
+        \hline
+        \textbf{Statistic} & \textbf{LightFM} & \textbf{SVD} & \textbf{Most Popular} \\ \hline
+        Precision          & 0.3660           & 0.2020       & 0.3010                \\
+        Recall@5           & 0.0942           & 0.0520       & 0.0766                \\
+        nDCG@5             & 0.6232           & 0.4762       & 0.5903                \\
+        Relevance Score    & 2.76             & 1.59         & 2.32                  \\
+        Diversity          & 0.03             & 0.01         & 0.01                  \\
+        Novelty            & 5.89             & 8.52         & 5.70                  \\
+        Fairness           & 0.60             & 0.72         & 0.66                  \\
+        Serendipity        & 4.51             & 2.80         & 3.63                  \\ \hline
+    \end{tabular}
+    \caption{Random groups recomendation results}
+    \label{tab:resultados-modelos}
+\end{table}
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{|l|l|l|l|}
+        \hline
+        \textbf{Statistic} & \textbf{LightFM} & \textbf{SVD} & \textbf{Most Popular} \\ \hline
+        Precision          & 0.3350           & 0.2040       & 0.3240                \\
+        Recall@5           & 0.0821           & 0.0498       & 0.0749                \\
+        nDCG@5             & 0.5722           & 0.4736       & 0.5474                \\
+        Relevance Score    & 2.57             & 1.62         & 2.48                  \\
+        Diversity          & 0.04             & 0.01         & 0.01                  \\
+        Novelty            & 5.99             & 8.60         & 5.70                  \\
+        Fairness           & 0.63             & 0.72         & 0.64                  \\
+        Serendipity        & 4.61             & 2.75         & 4.06                  \\ \hline
+    \end{tabular}
+    \caption{Similar groups recommendation results}
+    \label{tab:resultados-modelos-similar}
+\end{table}
 

### 2024-12-09 paper v2/paper v2.tex
Resultados paper

@@
     \item Recall at k: This metric measures the proportion of relevant items that are recommended to the user.
-    \item F1 score: This metric is the harmonic mean of precision and recall, and is used to measure the overall performance of the model.
+    \item nDCG@k: This metric measures the normalized discounted cumulative gain at k, and is used to measure the quality of the recommendations.
+    \item Relevance Score: This metric evaluates how relevant the recommendations are to the user, providing an indication of the recommendation quality.
+    \item Diversity: This metric assesses the variety of the recommendations, ensuring a wide range of different items are suggested.
+    \item Novelty: This metric measures how new or unexpected the recommendations are to the user, contributing to the overall recommendation quality.
+    \item Fairness: This metric evaluates the fairness of the recommendations, ensuring that the recommendations are unbiased and equitable.
+    \item Serendipity: This metric measures the likelihood of the recommendations to pleasantly surprise the user, enhancing the user experience.
 \end{itemize}
 
+As is shown on table \ref{tab:resultados-modelos}, the LightFM model outperforms the other models in almost all metrics.
+This is to be expected, since its using not only user-item interactions but also metadata to make recommendations.
+The metadata recommendation allows for the model to recommend similiar items to two or more users that have similar preferences even if they have not interacted with any items in common.
+This is specially helpful when
+making groups recommendations, because it allows for greater diversity and novelty and creates a more personalized group experience.
+
+\newline
+One aspect of the recommendations that is worth mentioning is that the LightFM model is capable of superior results with more diversity.
+This means that all groups will get recommended different games, but they are tailored to the preferences of each user on a group.
+This in hand increases the novelty and the serendipity of the recommendations, creating more unique experiences for the users.
+
+\newline
+For each different group formation, the results are the same.
+This is to be expected, since the group formation is based on the preferences of the users, and the recommendations are made based on the preferences of the group.
+Not on the preferences of the individual users.
+This is a good sign, beacuase it shows that the model does not overfit to any group formation and it is able to make recommendations based on the preferences of the group.
+
+\newline
+
+However, for new users, the results might not be as expected. The user needs to create user-item interactions in order to get proper recommendations.
 \begin{table}[]

### 2024-12-09 paper v2/paper v2.tex
Resultados paper

@@
 \begin{enumerate}
-    \item Simple average: The preferences of the group are calculated as the simple average of the preferences of the individual players.
-    \item Least misery: The preferences of the group are calculated as the minimum of the preferences of the individual players.
-    \item ... % asegurarse que esto está al día con lo que presentemos al final del trabajo
+    \item Simple average: The preferences of the group are calculated as the simple average of the preferences of the individual players based on a top-100 recommendation list prior to the group formation.
+This can be done with the following formula:
+    \begin{equation}
+        \text{avg\_score\_item} = \frac{\sum_{i=1}^{n} \text{user\_rating}_i}{n}
+    \end{equation}
+    with $n$ being the number of users in the group, and $u_i$ being the rating of the $i$-th user given an item.
+    % \item Least misery: The preferences of the group are calculated as the minimum of the preferences of the individual players.
+    % \item ... % asegurarse que esto está al día con lo que presentemos al final del trabajo
 \end{enumerate}
+
+Given the method explained above, we assign the score to each of the items in the top-100 recommedation list of each user on a group.
+Then we create a new list of items combining the top-100 recommendation lists of each user on a group, and we extract the top-n items with the highest average scores.
+Note that for this to work properly, if two users have a matching item in their top-100 recommendation list, the item will only be included as the sum of the average rankings by the users in the group.
 
@@
 \end{itemize}
+
+For each metric, we used a cumulative avarage of the results of 300 different group formations, on which we applied the common formulas for each metric.
+Meaning that for each group, we calculated all of the metrics mentioned above, then we added them and divided by the number of groups.
+\newline
 

### 2024-12-09 paper v2/paper v2.tex
feat: future work

@@
 
+For this project we will use the LightFM library\footnote{\cite{DBLP:conf/recsys/Kula15}} to make the recommendations, which is a Python implementation of a factorization machine for collaborative filtering.
+
+We will also use SVD models from the scikit-learn library\footnote{\cite{scikit-learn}} to compare the results of our model with a more traditional approach, and the most popular model as a baseline.
+SVD was selected above other models like k-nearest due to its usage as industry standard\footnote{\cite{Roy2022}} and due to taking less time executing.
+
 \subsection{Data preprocessing}
@@
 
-To evaluate the results of our model, we will compare it to most popular as a sanity check, to iknn applied to the same dataset and to SVD applied to the same dataset.
+To evaluate the results of our model, we will compare it to most popular as a sanity check and to SVD applied to the same dataset and groups.
 % Agregar AGREE si logramos hacerlo funcionar ahora que tenemos menos datos.
@@
 \begin{itemize}
-    \item Precision at k: This metric measures the proportion of recommended items that are relevant to the user, where relevance is defined as the items that the user has interacted with.
-    \item Recall at k: This metric measures the proportion of relevant items that are recommended to the user.
+    \item Precision@k: This metric measures the proportion of recommended items that are relevant to the user, where relevance is defined as the items that the user has interacted with.
+    \item Recall@k: This metric measures the proportion of relevant items that are recommended to the user.
     \item nDCG@k: This metric measures the normalized discounted cumulative gain at k, and is used to measure the quality of the recommendations.
@@
 
-\newline
 One aspect of the recommendations that is worth mentioning is that the LightFM model is capable of superior results with more diversity.
@@
 
-\newline
-For each different group formation, the results are the same.
+For each different group formation, the results are mostly the same.
 This is to be expected, since the group formation is based on the preferences of the users, and the recommendations are made based on the preferences of the group.
@@
 
-\newline
+However, for new users, the results might not be as expected.
+The user needs to create user-item interactions in order to get proper recommendations.
+Untill then, the model should be able to recommend the most popular items, which are the most likely to be relevant to the user according to the very low information the model has.
 
-However, for new users, the results might not be as expected. The user needs to create user-item interactions in order to get proper recommendations.
 \begin{table}[]
@@
         \textbf{Statistic} & \textbf{LightFM} & \textbf{SVD} & \textbf{Most Popular} \\ \hline
-        Precision          & 0.3660           & 0.2020       & 0.3010                \\
+        Precision@5        & 0.3660           & 0.2020       & 0.3010                \\
         Recall@5           & 0.0942           & 0.0520       & 0.0766                \\
@@
         \textbf{Statistic} & \textbf{LightFM} & \textbf{SVD} & \textbf{Most Popular} \\ \hline
-        Precision          & 0.3350           & 0.2040       & 0.3240                \\
+        Precision@5        & 0.3350           & 0.2040       & 0.3240                \\
         Recall@5           & 0.0821           & 0.0498       & 0.0749                \\
@@
 
+\section{Future work}
+
+Since this project was done with a reduced dataset, the obvious step would be to scale it up to the full dataset or at least a bigger portion of it with improved computation power.
+This would allow for more accurate results and a more robust set of models.
+
+Another step would be to improve the group formation algorithm, as it is currently based on the preferences of the users, and not on the characteristics of the games they play (it does not use the available metadata).
+This could be done by using a clustering algorithm to group the games based on their characteristics, and then grouping the users based on the games they have interacted with.
+Future work could also use a hybrid approach, where part of how the groups are formed is random and another part is based on the similarity of the users.
+
+Additionally, a notable drawback of this work is the lack of inclusion of the baseline "AGREE" recommended by the course instructors of IIC3633.
+Since the dataset selection process was finalized late in the development of this work, this implementation was not completed.
+
+Finally, there is exploration to do regarding the use of the least misery and dictatorship aggregation functions, as well as the use of other aggregation functions that could be more suitable for group recommendations.
+This could be done by comparing the results of the different aggregation functions and selecting the one that provides the best recommendations.
+This barely didnt make it into the project, and could easily be implemented in the future.
+
 \bibliography{main}

### 2024-12-09 metrics.py
Update metrics.py
Errores de formateo y rutas arreglados

@@ -5,7 +5,7 @@ import pandas as pd
 def parse_group_data(groups_df):
     parsed_data = []
     
-    for _, row in groups_df.head(300).iterrows():
+    for _, row in groups_df.head(200).iterrows():
         group_members = ast.literal_eval(row['group_members'])
         recommendations = ast.literal_eval(row['recommendations'])
         parsed_data.append((group_members, recommendations))
@@ -143,13 +143,24 @@ def calculate_group_metrics(parsed_data, train_df, k=5, threshold=5):
     }
     return avg_metrics
 
-train_df = pd.read_csv('train.csv')
-# df = pd.read_csv('RecSysProyectoFinal/random_group_recommendations.csv', sep=';')
-# df.to_csv('RecSysProyectoFinal/random_group_recommendations.csv', index=False, quoting=1)
-groups_df = pd.read_csv('RecSysProyectoFinal/random_group_recommendations.csv')
+test_df = pd.read_csv('recommendations/lastsample/test.csv')
+groups_df = pd.read_csv('recommendations/lastsample/random_group_recommendations.csv')
 parsed_data = parse_group_data(groups_df)
-metrics = calculate_group_metrics(parsed_data, train_df, k=5)
-print("Random LightFM")
+metrics = calculate_group_metrics(parsed_data, test_df, k=5)
+print("\nRandom LightFM")
+print(f"Precision: {metrics['average_precision']:.6f}")
+print(f"Recall: {metrics['average_recall']:.6f}")
+print(f"nDCG: {metrics['average_ndcg']:.6f}")
+print(f"Relevance Score: {metrics['average_relevance']:.2f}")
+print(f"Diversity: {metrics['average_diversity']:.2f}")
+print(f"Novelty: {metrics['average_novelty']:.2f}")
+print(f"Fairness: {metrics['average_fairness']:.2f}")
+print(f"Serendipity: {metrics['average_serendipity']:.2f}")
+
+groups_df = pd.read_csv('recommendations/lastsample/similar_group_recommendations.csv')
+parsed_data = parse_group_data(groups_df)
+metrics = calculate_group_metrics(parsed_data, test_df, k=5)
+print("\nSimilar LightFM")
 print(f"Precision: {metrics['average_precision']:.6f}")
 print(f"Recall: {metrics['average_recall']:.6f}")
[…]

### 2024-12-09 paper v2/paper v2.tex
fix: no \newline !!!

--- 2024-12-09 paper v2/paper v2.tex
@@
 \usepackage{comment}
+\usepackage{amsmath}
 
@@
 
-\newline
 For each different group formation, the results are the same.
@@
 This is a good sign, beacuase it shows that the model does not overfit to any group formation and it is able to make recommendations based on the preferences of the group.
-
-\newline
 

### 2024-12-09 metrics.py
metrics.py parametrizado

@@ -143,8 +143,11 @@ def calculate_group_metrics(parsed_data, train_df, k=5, threshold=5):
     }
     return avg_metrics
 
-test_df = pd.read_csv('recommendations/lastsample/test.csv')
-groups_df = pd.read_csv('recommendations/lastsample/random_group_recommendations.csv')
+
+sampling_method = "itemsaaaample"   # ["itemsaaaample", "lastsaaaample"] escoger una
+
+test_df = pd.read_csv(f'recommendations/{sampling_method}/test.csv')
+groups_df = pd.read_csv(f'recommendations/{sampling_method}/random_group_recommendations.csv')
 parsed_data = parse_group_data(groups_df)
 metrics = calculate_group_metrics(parsed_data, test_df, k=5)
 print("\nRandom LightFM")
@@ -157,7 +160,7 @@ print(f"Novelty: {metrics['average_novelty']:.2f}")
 print(f"Fairness: {metrics['average_fairness']:.2f}")
 print(f"Serendipity: {metrics['average_serendipity']:.2f}")
 
-groups_df = pd.read_csv('recommendations/lastsample/similar_group_recommendations.csv')
+groups_df = pd.read_csv(f'recommendations/{sampling_method}/similar_group_recommendations.csv')
 parsed_data = parse_group_data(groups_df)
 metrics = calculate_group_metrics(parsed_data, test_df, k=5)
 print("\nSimilar LightFM")
@@ -171,7 +174,7 @@ print(f"Fairness: {metrics['average_fairness']:.2f}")
 print(f"Serendipity: {metrics['average_serendipity']:.2f}")
 
 
-groups_df = pd.read_csv('recommendations/lastsample/svd_random_group_recommendations.csv')
+groups_df = pd.read_csv(f'recommendations/{sampling_method}/svd_random_group_recommendations.csv')
 parsed_data = parse_group_data(groups_df)
 metrics = calculate_group_metrics(parsed_data, test_df, k=5)
 print("\nRandom SVD")
@@ -184,7 +187,7 @@ print(f"Novelty: {metrics['average_novelty']:.2f}")
 print(f"Fairness: {metrics['average_fairness']:.2f}")
 print(f"Serendipity: {metrics['average_serendipity']:.2f}")
 
-groups_df = pd.read_csv('recommendations/lastsample/svd_similar_group_recommendations.csv')
[…]

### 2024-12-09 paper v2/paper v2.tex
fix: gitignore
feat: conclusión

--- 2024-12-09 paper v2/paper v2.tex
@@
 
-\section{Sensitivity Analysis}
-
 \section{Conclusions}
 
-% You can use footnotes\footnote{Footnotes should be complete sentences.} to provide readers with additional information about a topic without interrupting the flow of the paper.
+In this work, we presented a novel approach to group-based board game recommendations by leveraging both user ratings and game metadata.
+Our system outperforms traditional recommendation models, such matrix factorization (SVD), in terms of key metrics like Precision, Recall, nDCG, Diversity, and Serendipity.
+By incorporating metadata — including attributes like game mechanics, themes, and complexity — our model provides more personalized and diverse recommendations that cater to the collective preferences of the group, rather than relying solely on individual user interactions.
 
-% \begin{figure}[ht]
-% \vskip 0.2in
-% \begin{center}
-% \centerline{\includegraphics[width=\columnwidth]{icml_numpapers}}
-% \caption{Historical locations and number of accepted papers for International
-% Machine Learning Conferences (ICML 1993 -- ICML 2008) and International
-% Workshops on Machine Learning (ML 1988 -- ML 1992). At the time this figure was
-% produced, the number of accepted papers for ICML 2008 was unknown and instead
-% estimated.}
-% \label{icml-historical}
-% \end{center}
-% \vskip -0.2in
-% \end{figure}
-
-% \section{Figures}
-
-% You may want to include figures in the paper to illustrate your approach and results.
-Such artwork should be centered, legible, and separated from the text.
-Lines should be dark and at least 0.5~points thick for purposes of reproduction, and text should not appear on a gray background.
-
-% Label all distinct components of each figure.
-If the figure takes the form of a graph, then give a name for each axis and include a legend that briefly describes each curve.
-Do not include a title inside the figure; instead, the caption should serve this function.
-
-%Number figures sequentially, placing the figure number and caption \emph{after} the graphics, with at least 0.1~inches of space before
-
-%You may float figures to the top or bottom of a column, and you may set wide figures across both columns (use the environment \texttt{figure*} in \LaTeX).
-Always place two-column figures at the top or bottom of the page.
-
-% Ejemplo de algoritmo (creo que no usaremos esto)
-% \begin{algorithm}[tb]
-%    \caption{Bubble Sort}
-%    \label{alg:example}
-% \begin{algorithmic}
-%    \STATE {\bfseries Input:} data $x_i$, size $m$
-%    \REPEAT
-%    \STATE Initialize $noChange = true$.
-%    \FOR{$i=1$ {\bfseries to} $m-1$}
-%    \IF{$x_i > x_{i+1}$}
-%    \STATE Swap $x_i$ and $x_{i+1}$
-%    \STATE $noChange = false$
-%    \ENDIF
-%    \ENDFOR
-%    \UNTIL{$noChange$ is $true$}
-% \end{algorithmic}
-% \end{algorithm}
-
-% Ejemplo de tabla
-% \begin{table}[t]
-% \caption{Classification accuracies for naive Bayes and flexible
-% Bayes on various data sets.}
-% \label{sample-table}
-% \vskip 0.15in
-% \begin{center}
-% \begin{small}
-% \begin{sc}
-% \begin{tabular}{lcccr}
-% \toprule
-% Data set & Naive & Flexible & Better? \\
-% \midrule
-% Breast    & 95.9$\pm$ 0.2& 96.7$\pm$ 0.2& $\surd$ \\
-% Cleveland & 83.3$\pm$ 0.6& 80.0$\pm$ 0.6& $\times$\\
-% Glass2    & 61.9$\pm$ 1.4& 83.8$\pm$ 0.7& $\surd$ \\
-% Credit    & 74.8$\pm$ 0.5& 78.3$\pm$ 0.6&         \\
-% Horse     & 73.3$\pm$ 0.9& 69.7$\pm$ 1.0& $\times$\\
-% Meta      & 67.1$\pm$ 0.6& 76.5$\pm$ 0.5& $\surd$ \\
-% Pima      & 75.1$\pm$ 0.6& 73.9$\pm$ 0.5&         \\
-% Vehicle   & 44.9$\pm$ 0.6& 61.5$\pm$ 0.4& $\surd$ \\
-% \bottomrule
-% \end{tabular}
-% \end{sc}
-% \end{small}
-% \end{center}
-% \vskip -0.1in
-% \end{table}
-
-% \subsection{Citations and References}
-
-% Please use APA reference format regardless of your formatter
-% or word processor. If you rely on the \LaTeX\/ bibliographic
-% facility, use \texttt{natbib.sty} and \texttt{icml2021.bst}
-% included in the style-file package to obtain this format.
-
-% Citations within the text should include the authors' last names and
-% year. If the authors' names are included in the sentence, place only
-% the year in parentheses using yrcite.
-
-% Alphabetize references by the surnames of the first authors, with
-% single author entries preceding multiple author entries. Order
-% references for the same authors by year of publication, with the
-% earliest first. Make sure that each reference includes all relevant
-% information (e.g., page numbers).
-
-% Please put some effort into making references complete, presentable, and
-% consistent. If using bibtex, please protect capital letters of names and
-% abbreviations in titles, for example, use \{B\}ayesian or \{L\}ipschitz
-% in your .bib file.
-
-% In the unusual situation where you want a paper to appear in the
-% references without citing it in the main text, use \nocite
-% \nocite{}
+The results highlight the value of metadata in enhancing recommendation accuracy, ensuring greater variety and novelty in the suggested games, and improving the overall user experience.
+Our method not only improves the quality of recommendations but also addresses the challenge of recommending games for groups with diverse preferences.
+The robustness and scalability of the proposed system indicate its potential for application in other domains requiring group-based recommendations, such as tourism, movies, or event planning.
 

### 2024-12-09 paper v2/paper v2.tex

@@
 
-In this work, we presented a novel approach to group-based board game recommendations by leveraging both user ratings and game metadata.
+In this work, we proposed a group recommendation system for board games using metadata from the Board Game Geek website.
+The system uses a LightFM model to make recommendations based on the preferences of the group.
+
 Our system outperforms traditional recommendation models, such matrix factorization (SVD), in terms of key metrics like Precision, Recall, nDCG, Diversity, and Serendipity.
@@
 
-The results highlight the value of metadata in enhancing recommendation accuracy, ensuring greater variety and novelty in the suggested games, and improving the overall user experience.
+The results are consistent across different group formations, showing that the model is able to make recommendations based on the preferences of the group, and not on the preferences of the individual users.
+This is a key advantage of our system, as it ensures that the recommendations are tailored to the group as a whole, rather than to individual users.
+
+The results also highlight the value of metadata in enhancing recommendation accuracy, ensuring greater variety and novelty in the suggested games, and improving the overall user experience.
 Our method not only improves the quality of recommendations but also addresses the challenge of recommending games for groups with diverse preferences.
-The robustness and scalability of the proposed system indicate its potential for application in other domains requiring group-based recommendations, such as tourism, movies, or event planning.
+
+This work shows that metadata-based group recommendation systems can be an effective way to make recommendations for board games, especially with low computational power available.
+The robustness and scalability of the proposed system indicate its potential for application in other domains requiring group-based recommendations, such as tourism, movies, or event planning, assuming the metadata is available.
 

### 2024-12-09 metrics.py
Update metrics.py
sampling_method

@@ -144,7 +144,7 @@ def calculate_group_metrics(parsed_data, train_df, k=5, threshold=5):
     return avg_metrics
 
 
-sampling_method = "itemsaaaample"   # ["itemsaaaample", "lastsaaaample"] escoger una
+sampling_method = "itemsample"   # ["itemsample", "lastsample"] escoger una
 
 test_df = pd.read_csv(f'recommendations/{sampling_method}/test.csv')
 groups_df = pd.read_csv(f'recommendations/{sampling_method}/random_group_recommendations.csv')

### 2024-12-09 paper v2/paper v2.tex
fix final del paper

@@
 
-We propose a refined version of that last system, using metadata-centric clustering to group board games based on their characteristics, and then recommending games to groups of players based on their summed preferences used the functions detailed on the previous work.
+We propose a refined version of that last system, using metadata-centric clustering to group board games based on their characteristics, and then recommending games to groups of players based on their summed preferences using the functions detailed on the previous work.
 This system is designed to be more accuarate than the previous one, as it takes into account the characteristics of the games, and not just the ratings of the users.
@@
     \item \texttt{subcategories.csv}: Contains the subcategories of each game in binary flag format.
-    \item \texttt{artists\_reduced.csv}: Contains information about which artist created a certain game in binary flag format. Only artists with more than 3 games created will be considered.
-    \item \texttt{designer\_reduced.csv}: Contains information about which designer designed a certain game in binary flag format. Only designers with more than 3 games created are considered.
-    \item \texttt{publishers\_reduced.csv}: Contains information about the companies that sell these games in binary flag format. Only companies that sell more than 3 games are considered.
+    \item \texttt{artists\_reduced.csv}: Contains information about which artist created a certain game in binary flag format.
+    \item \texttt{designer\_reduced.csv}: Contains information about which designer designed a certain game in binary flag format.
+    \item \texttt{publishers\_reduced.csv}: Contains information about which companies sell each game in binary flag format.
 
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{publishYear.png}
+    \includegraphics[width=0.9\linewidth]{publishYear.png}
     \caption{Amount of games (y axis) by year of publication(x axis)}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{distribucionRatings.png}
+    \includegraphics[width=0.9\linewidth]{distribucionRatings.png}
     \caption{Games rating distribution (1-10)}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{tematicasComunes.png}
+    \includegraphics[width=0.9\linewidth]{tematicasComunes.png}
     \caption{Most common game themes on the dataset}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{mecanicasComunes.png}
+    \includegraphics[width=0.9\linewidth]{mecanicasComunes.png}
     \caption{Most common game mechanics on the dataset}
@@
 
-With that in mind, we will reduce the amount of games to 1500, from the original $>$10k by taking 375 games from each class.
+With that in mind, we will reduce the amount of games to 1500, from the original 22k by taking 375 games from each class.
 
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{ReviewsPorJuego.png}
+    \includegraphics[width=0.9\linewidth]{ReviewsPorJuego.png}
     \caption{Amount of reviews per game}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{reviewsPorUsuario.png}
+    \includegraphics[width=0.9\linewidth]{reviewsPorUsuario.png}
     \caption{Amount of reviews per user (full dataset)}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{reviewsPorUsuario2.png}
+    \includegraphics[width=0.9\linewidth]{reviewsPorUsuario2.png}
     \caption{Amount of reviews per user (1500 items)}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{reviewsPorUsuarioFinal.png}
+    \includegraphics[width=0.9\linewidth]{reviewsPorUsuarioFinal.png}
     \caption{Amount of reviews per user (after full reduction)}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{ReviewsPorJuegoFinal.png}
+    \includegraphics[width=0.9\linewidth]{ReviewsPorJuegoFinal.png}
     \caption{Amount of reviews per game (after full reduction)}
@@
 The second stage consists of grouping the users based on their characteristics by using a correlation matrix.
-The function used is an adaptation of the one used in the "" work, which is based on the cosine similarity between the users.
-The function is as follows:
+The function used is an adaptation of the one used in the work of \cite{group_recommenders_repo}.
+This function is used to calculate the similarity between users based on their ratings, and then group them based on that similarity.
 
-% Arreglar función
-\begin{equation}
-    %\text{similarity}(u, v) = \frac{\sum_{i=1}^{n} u_i \cdot v_i}{\sqrt{\sum_{i=1}^{n} u_i^2} \cdot \sqrt{\sum_{i=1}^{n} v_i^2}}
-\end{equation}
+To assemble groups by similarity, we use the Pearson product-moment correlation coefficients matrix.
+In order to do this, a matrix with users by rows and items by columns is built, in which the entry $i,j$ is the rating that the user $i$ gave to item $j$, every item that is not rated is rated as 0.
+
+From that matrix, we calculate the coefficient matrix where the entry $i,j$ is $$\frac{C_{ij}}{\sqrt{C_{ii}C_{jj}}}$$ with $C_{ij}$  the covariance between user $i$ and user $j$.
+
+The groups are then formed by selecting the users with the highest correlation coefficients.
+The number of groups is determined by the user, and the number of users per group is also determined by the user.
+For this work, we will use as many groups as possible with 4 users each (as it is the most common maximum of players for a given game in this dataset, as seen on figure \ref{fig:output}).
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.9\linewidth]{output.png}
+    \caption{Most common number of max players}
+    \label{fig:output}
+\end{figure}
 
@@
 For each metric, we used a cumulative avarage of the results of 300 different group formations, on which we applied the common formulas for each metric.
-Meaning that for each group, we calculated all of the metrics mentioned above, then we added them and divided by the number of groups.
-\newline
+Meaning that, for each group, we calculated all of the metrics mentioned above, then we added them and divided by the number of groups.
 
-As is shown on table \ref{tab:resultados-modelos}, the LightFM model outperforms the other models in almost all metrics.
-This is to be expected, since its using not only user-item interactions but also metadata to make recommendations.
+As is shown on tables \ref{tab:resultados-modelos} and \ref{tab:resultados-modelos-similar}, the LightFM model outperforms the other models in almost all metrics.
+This is what we where looking for in this work, that it gives better metrics since its using not only user-item interactions but also metadata to make recommendations.
 The metadata recommendation allows for the model to recommend similiar items to two or more users that have similar preferences even if they have not interacted with any items in common.
@@
 Our system outperforms traditional recommendation models, such matrix factorization (SVD), in terms of key metrics like Precision, Recall, nDCG, Diversity, and Serendipity.
-By incorporating metadata — including attributes like game mechanics, themes, and complexity — our model provides more personalized and diverse recommendations that cater to the collective preferences of the group, rather than relying solely on individual user interactions.
+By incorporating metadata — including attributes like game mechanics, themes, and complexity — our model provides more personalized and diverse recommendations.
 
